---
title: "Harvest Rate Titration"
author: "Colin Dassow"
output: bookdown::pdf_document2
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, cache = T, tidy = TRUE, tidy.opts = list(width.cutoff = 60))

rm(list=ls())
library(wdnr.fmdb)# contact the author about the use of this package which is not currently available on CRAN
library(tinytex)
library(tidyverse)
library(ggpubr)
library(knitr)
library(BayesianTools)
library(moments)

# set working directory to location of local copy of git repo
setwd("C:/Users/dassocju/Documents/OAS_git_repos/Dassow.et.al._Creel_Efficiencies")

#reading in objects that take a long time to produce to improve rendering speed of this document
call=readRDS("creelDataSet_all.RData")
# unlist, etc. to get back to individual dfs, filtering to completed trips only
cserv=call[[1]]
cvis=call[[2]]
ccou=call[[3]]
cint=call[[4]]; cint=cint[cint$trip.completed=='y',]
cfish=call[[5]]; cfish=cfish[cfish$trip.complete=='y',]
cfish.i=call[[6]]; cfish.i=cfish.i[cfish.i$trip.complete.flag=='y',] 


# making list of CTWI lakes with creels
lchar=get_fmdb_lakechar()
ctwiWBIC=lchar[lchar$trtystat==1 & !is.na(lchar$trtystat),] # picking out just lakes in CTWI
ctwiWBIC.creel=ctwiWBIC$wbic[ctwiWBIC$wbic%in%cserv$wbic] # wbics in CTWI that have been creeled

```

# WDNR Creel Titration

## Current Creel Data as Baseline

DNR has amassed a large amount of inland creel information over many decades (`r range(cserv$year)`) across many waterbodies (n=`r length(unique(cserv$wbic))`).
Using this data and functionality of `wdnr.fmdb` it is possible to calculate effort, catch, harvest, and harvest rate for multiple species for each unique creel survey conducted.
This information can be further grouped by month, season, day type, etc. to provide insight into important temporal patterns in the fisheries metrics of interest.
These calculations are accomplished fairly easily using the following code:

```{r creel-data-retreival, eval=FALSE, echo=T}

# reading in DNR creel data

cserv=get_creel_surveys()
cvis=get_creel_visits()
ccou=get_creel_counts()
cint=get_creel_int_party()
cfish=get_creel_fish_data()
```

```{r baseline-fisheries-metrics, echo=T}

ceff=calc_creel_effort(creel_count_data=ccou, creel_int_data=cint) 

charv=calc_creel_harvest(creel_count_data = ccou, creel_int_data = cint,creel_fish_data = cfish) 

charvR=calc_creel_harvest_rates(creel_fish_data=cfish)

```

```{r u-rate-calcs}

## CALC EXPLOITATION RATES ####
# keeping the same code as the original exploitation rate analysis for consistency
tagdat=read.csv("tags_and_marks.csv") # marking data from fmdb since the package function URLs didn't seem to be working this was downloaded manually from the database here: https://dnr.wisconsin.gov/topic/Fishing/data/infosystem.html
tagdat$County=standardize_county_names(tagdat$County)
tagdat$Gear=standardize_string(tagdat$Gear)
tagdat$Waterbody.Name=standardize_waterbody_names(tagdat$Waterbody.Name)

fndat=tagdat%>% # this is just in case we care what type of mark is there, but I'm going to overwrite this for now and just get total number marked.
  filter(WBIC%in%ctwiWBIC.creel & Species.Code=="X22" & Gear=="fyke_net", !is.na(Mark.Given) & Mark.Given!="")%>%
  group_by(WBIC, Waterbody.Name,Survey.Year,Survey.Begin.Date,Survey.End.Date,Survey.Seq.No,Mark.Given)%>%
  summarise(nFish=sum(Number.of.Fish))

# now summing across all types of marks
fndat=tagdat%>% 
  filter(WBIC%in%ctwiWBIC.creel & Species.Code=="X22" & Gear=="fyke_net", !is.na(Mark.Given) & Mark.Given!="")%>%
  group_by(WBIC, Waterbody.Name,Survey.Year,Survey.Begin.Date,Survey.End.Date,Survey.Seq.No)%>%
  summarise(nFish=sum(Number.of.Fish))%>%
  mutate(Survey.Begin.Date=lubridate::mdy(Survey.Begin.Date),
         Survey.End.Date=lubridate::mdy(Survey.End.Date),
         begin.md=format(Survey.Begin.Date, "%m-%d"),
         end.md=format(Survey.End.Date, "%m-%d"))%>%
  filter(begin.md<"06-01") # getting rid of any marking surveys that may have taken place well after fishing season opened.

fdat=fndat%>% # now that I have only surveys I want, pooling across start and end dates within a year to create an annual total number marked for comparison to creel data
  group_by(WBIC, Waterbody.Name, Survey.Year)%>%
  summarise(nFN.marked=sum(nFish))

# table of marks found during creel surveys
crRecap=cfish.i%>%
  filter(species.code=="X22" & wbic%in%ctwiWBIC.creel)%>%
  mutate(month=month(sample.date),
         daytype=ifelse(wday(sample.date)%in%c(1,7),"weekend","weekday"))%>%
  group_by(wbic, year,survey.seq.no,survey.begin.date,survey.end.date, mark.found,month,daytype)%>%
  summarise(nfish=sum(fish.count)) # number of fish of each mark typed returned in that creel
crRecap=crRecap%>% # now throwing out mark type and just taking overall number of marks recapped that month for that survey.
  group_by(wbic, year, survey.seq.no, survey.begin.date, survey.end.date,month,daytype)%>%
  summarise(n.marks.recapped=sum(nfish[!is.na(mark.found)]),
            totalFishExamined=sum(nfish))

# spp.harvest is the column that has the estimate of the harvest for that species in that strata
harvestEstimates=charv%>%
  filter(species.code=="X22" & wbic%in%ctwiWBIC.creel)%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest,harvest,total.harvest,total.spp.harvest)%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(harvAcre=spp.harvest/lake.area)

# combining harvest estimates with marked proportion estimate to get total number of marks harvested

markHarvEst=crRecap%>%
  left_join(harvestEstimates)%>%
  mutate(markHarvEstimate=(n.marks.recapped/totalFishExamined)*spp.harvest)

ang.exp=markHarvEst%>%
  left_join(fdat, by=c("wbic"="WBIC","year"="Survey.Year"))%>%
  mutate(exp.rate=markHarvEstimate/nFN.marked) # 
naExps=ang.exp[is.na(ang.exp$exp.rate),] # looking at the lake-years without data to see if there's a problem with my code or just missing data
# looks like it's either cases were no fish were marked in spring fyking for that creel survey or if for a specific strata no walleye were harvested or no marked walleye were harvested, either of those will throw an NA in the exploitation rate calculation

# there are 6 observation where the number of marks returned was higher than what was marked, I'm throwing these out.
ang.exp=ang.exp[!is.na(ang.exp$exp.rate) & ang.exp$exp.rate<1.0,]

ang.exp=ang.exp%>%
  group_by(survey.seq.no)%>%
  summarise(meanU=mean(exp.rate,na.rm = T),
            exp.rate.sd=sd(exp.rate,na.rm = T))

# now I have exploitation rates and as a result harvest rates too since they get calculated along the way

```

```{r thinning data}
## CREATING REDUCED DATASETS ####
# next I want to make a few scenarios with 'reduced' data, a 'noWinter' scenario, 'May-August' scenario, and 3 % removals (0.25 weekdays, 0.5 weekdays, 0.5 weekend)

ceff.ct=ceff%>% # effort data frame
  filter(wbic%in%ctwiWBIC.creel)%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(annual.eff.acre=total.effort/lake.area)

ccatch.ct=charv%>% # catch data frame
  filter(wbic%in%ctwiWBIC.creel, species=='walleye')%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(annual.catch.acre=catch/lake.area) # catch is the estimated total catch for that strata (total.effort * catch rate), while 'total.catch' is the actually number of fish in the interviews for that strata

ifish=cfish.i%>% # making a data frame to add my groupings too so I can leave cfish.i alone
  filter(species.code=="X22" & wbic%in%ctwiWBIC.creel)%>%
  mutate(daytype=ifelse(wday(sample.date)%in%c(1,7),"weekend","weekday"),
         month=month(sample.date),
         season=ifelse(month%in%4:10,"openwater","winter"))

iharv=charv%>% # making a separate data frame to add groupings to so I can leave charv alone
  filter(species.code=="X22" & wbic%in%ctwiWBIC.creel)%>%
  mutate(season=ifelse(month%in%4:10,"openwater","winter"))

iharvR=charvR%>% # making a separate data frame to add groupings to so I can leave charvR alone
  filter(species.code=="X22" & wbic%in%ctwiWBIC.creel)%>%
  mutate(season=ifelse(month%in%4:10,"openwater","winter"))

# now making a reduced dataframe for each scenario

###### NO WINTER ####

eff.nw=ceff.ct%>%
  filter(season!='winter')

catch.nw=ccatch.ct%>%
  filter(month%in%ceff$month[ceff$season!='winter'])

ifish.nw=ifish%>%
  filter(season=="openwater")%>%
  group_by(wbic, year,survey.seq.no,survey.begin.date,survey.end.date, mark.found,month,daytype)%>%
  summarise(nfish=sum(fish.count)) # number of fish of each mark typed returned in that creel

ifish.nw=ifish.nw%>% # now throwing out mark type and just taking overall number of marks recapped that month for that survey.
  group_by(wbic, year, survey.seq.no, survey.begin.date, survey.end.date,month,daytype)%>%
  summarise(n.marks.recapped=sum(nfish[!is.na(mark.found)]),
            totalFishExamined=sum(nfish))

# spp.harvest is the column that has the estimate of the harvest for that species in that strata
harvestEstimates.nw=iharv%>%
  filter(season=="openwater")%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest,harvest,total.harvest,total.spp.harvest)%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(harvAcre=spp.harvest/lake.area)

harvestR.nw=iharvR%>%
  filter(season=="openwater")%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest.rate,harvest.rate)

# combining harvest estimates with marked proportion estimate to get total number of marks harvested

markHarvEst.nw=ifish.nw%>%
  left_join(harvestEstimates.nw)%>%
  mutate(markHarvEstimate=(n.marks.recapped/totalFishExamined)*spp.harvest)

ang.exp.nw=markHarvEst.nw%>%
  left_join(fdat, by=c("wbic"="WBIC","year"="Survey.Year"))%>%
  mutate(exp.rate=markHarvEstimate/nFN.marked)

# Throwing out NA u's and those >1 (NAs from lack of walleye harvest per strata or missing FN marks, >1's is 6 observations where # marks in creel exceeds # marks at large)
ang.exp.nw=ang.exp.nw[!is.na(ang.exp.nw$exp.rate) & ang.exp.nw$exp.rate<1.0,]
# creating survey-level exploitation rates
ang.exp.nw=ang.exp.nw%>%
  group_by(survey.seq.no)%>%
  summarise(meanU=mean(exp.rate,na.rm = T),
            exp.rate.sd=sd(exp.rate,na.rm = T))

###### MAY-AUG ####

eff.ma=ceff.ct%>%
  filter(month%in%c(5:8))

catch.ma=ccatch.ct%>%
  filter(month%in%c(5:8))

ifish.ma=ifish%>%
  filter(month%in%c(5:8))%>%
  group_by(wbic, year,survey.seq.no,survey.begin.date,survey.end.date, mark.found,month,daytype)%>%
  summarise(nfish=sum(fish.count)) # number of fish of each mark typed returned in that creel

ifish.ma=ifish.ma%>% # now throwing out mark type and just taking overall number of marks recapped that month for that survey.
  group_by(wbic, year, survey.seq.no, survey.begin.date, survey.end.date,month,daytype)%>%
  summarise(n.marks.recapped=sum(nfish[!is.na(mark.found)]),
            totalFishExamined=sum(nfish))

# spp.harvest is the column that has the estimate of the harvest for that species in that strata
harvestEstimates.ma=iharv%>%
  filter(month%in%c(5:8))%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest,harvest,total.harvest,total.spp.harvest)%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(harvAcre=spp.harvest/lake.area)

harvestR.ma=iharvR%>%
  filter(month%in%c(5:8))%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest.rate,harvest.rate)

# combining harvest estimates with marked proportion estimate to get total number of marks harvested

markHarvEst.ma=ifish.ma%>%
  left_join(harvestEstimates.ma)%>%
  mutate(markHarvEstimate=(n.marks.recapped/totalFishExamined)*spp.harvest)

ang.exp.ma=markHarvEst.ma%>%
  left_join(fdat, by=c("wbic"="WBIC","year"="Survey.Year"))%>%
  mutate(exp.rate=markHarvEstimate/nFN.marked)

# Throwing out NA u's and those >1 (NAs from lack of walleye harvest per strata or missing FN marks, >1's is 6 observations where # marks in creel exceeds # marks at large)
ang.exp.ma=ang.exp.ma[!is.na(ang.exp.ma$exp.rate) & ang.exp.ma$exp.rate<1.0,]
# creating survey-level exploitation rates
ang.exp.ma=ang.exp.ma%>%
  group_by(survey.seq.no)%>%
  summarise(meanU=mean(exp.rate,na.rm = T),
            exp.rate.sd=sd(exp.rate,na.rm = T))

# for each of these % reductions I'm removing data based on visit fish seq no to simulate what would happen if that creel clear visit to the lake were removed.
###### 25% WEEKDAYS ####

rm.fish.seq.no=NA
rm.visit.fish.seq.no=NA
ifish.surv=unique(ifish$survey.seq.no)

set.seed(3)
for(i in 1:length(ifish.surv)){ # survey loop
  full=ifish[ifish$survey.seq.no==ifish.surv[i] & ifish$daytype=="weekday",]
  ms=unique(full$month)
  if(nrow(full)>0){
    for(m in 1:length(ms)){ # month loop
      visits=unique(full$visit.fish.seq.no[full$month==ms[m]])
      visit.rm=visits[c(round(runif(round(0.25*length(visits)),min=1,max = length(visits))))]
      t.rm=full$fish.data.seq.no[full$visit.fish.seq.no%in%visit.rm]# fish seq nos to remove that associated with the vist seq nos selected for removal
      
      rm.visit.fish.seq.no=c(rm.visit.fish.seq.no,visit.rm) # visit fish seq nos to use to reduce the interview, count, and fish data needed to get harvest estimates
      rm.fish.seq.no=c(rm.fish.seq.no,t.rm) # vector of fish to remove from individual fish data
      
    }
  }
}

ifish.25wd=ifish[!(ifish$fish.data.seq.no%in%rm.fish.seq.no),] # reduced individual fish data
iint.25wd=cint%>%
  filter(wbic%in%ctwiWBIC.creel & !(visit.fish.seq.no%in%rm.visit.fish.seq.no)) # reduced interview data
icou.25wd=ccou%>%
  filter(wbic%in%ctwiWBIC.creel & !(visit.fish.seq.no%in%rm.visit.fish.seq.no)) # reduced count data
ifishAg.25wd=cfish%>%
  filter(wbic%in%ctwiWBIC.creel & !(visit.fish.seq.no%in%rm.visit.fish.seq.no)) # reduced aggregate fish data

iharv.25wd=calc_creel_harvest(creel_count_data = icou.25wd,
                              creel_int_data = iint.25wd,
                              creel_fish_data = ifishAg.25wd) # harvest estimates from the reduced data
eff.25wd=calc_creel_effort(creel_count_data = icou.25wd, creel_int_data = iint.25wd)
eff.25wd=eff.25wd%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(annual.eff.acre=total.effort/lake.area)

catch.25wd=iharv.25wd%>%
  select(year,wbic,survey.seq.no, month, daytype, species, species.code, catch, spp.effort, spp.harvest, spp.catch)%>%
  filter(species=='walleye')%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(annual.catch.acre=catch/lake.area)

wd25=ifish.25wd%>%
  group_by(wbic, year,survey.seq.no,survey.begin.date,survey.end.date, mark.found,month,daytype)%>%
  summarise(nfish=sum(fish.count)) # number of fish of each mark typed returned in that creel

wd25=wd25%>% # now throwing out mark type and just taking overall number of marks recapped that month for that survey.
  group_by(wbic, year, survey.seq.no, survey.begin.date, survey.end.date,month,daytype)%>%
  summarise(n.marks.recapped=sum(nfish[!is.na(mark.found)]),
            totalFishExamined=sum(nfish))

# spp.harvest is the column that has the estimate of the harvest for that species in that strata
harvestEstimates.25wd=iharv.25wd%>%
  filter(species.code=="X22")%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest,harvest,total.harvest,total.spp.harvest)%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(harvAcre=spp.harvest/lake.area)

harvestR.25wd=calc_creel_harvest_rates(ifishAg.25wd)%>%
  filter(species.code=="X22")%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest.rate,harvest.rate)


# combining harvest estimates with marked proportion estimate to get total number of marks harvested

markHarvEst.25wd=wd25%>%
  left_join(harvestEstimates.25wd)%>%
  mutate(markHarvEstimate=(n.marks.recapped/totalFishExamined)*spp.harvest)

ang.exp.25wd=markHarvEst.25wd%>%
  left_join(fdat, by=c("wbic"="WBIC","year"="Survey.Year"))%>%
  mutate(exp.rate=markHarvEstimate/nFN.marked)

# Throwing out NA u's and those >1 (NAs from lack of walleye harvest per strata or missing FN marks, >1's is 6 observations where # marks in creel exceeds # marks at large)
ang.exp.25wd=ang.exp.25wd[!is.na(ang.exp.25wd$exp.rate) & ang.exp.25wd$exp.rate<1.0,]
# creating survey-level exploitation rates
ang.exp.25wd=ang.exp.25wd%>%
  group_by(survey.seq.no)%>%
  summarise(meanU=mean(exp.rate,na.rm = T),
            exp.rate.sd=sd(exp.rate,na.rm = T))


###### 50% WEEKDAYS ####

rm.fish.seq.no=NA
rm.visit.fish.seq.no=NA
ifish.surv=unique(ifish$survey.seq.no)

set.seed(3)
for(i in 1:length(ifish.surv)){ # survey loop
  full=ifish[ifish$survey.seq.no==ifish.surv[i] & ifish$daytype=="weekday",]
  ms=unique(full$month)
  if(nrow(full)>0){
    for(m in 1:length(ms)){ # month loop
      visits=unique(full$visit.fish.seq.no[full$month==ms[m]])
      visit.rm=visits[c(round(runif(round(0.50*length(visits)),min=1,max = length(visits))))]
      t.rm=full$fish.data.seq.no[full$visit.fish.seq.no%in%visit.rm]# fish seq nos to remove that associated with the visit seq nos selected for removal
      
      rm.visit.fish.seq.no=c(rm.visit.fish.seq.no,visit.rm) # visit fish seq nos to use to reduce the interview, count, and fish data needed to get harvest estimates
      rm.fish.seq.no=c(rm.fish.seq.no,t.rm) # vector of fish to remove from individual fish data
      
    }
  }
}

ifish.50wd=ifish[!(ifish$fish.data.seq.no%in%rm.fish.seq.no),] # reduced individual fish data
iint.50wd=cint%>%
  filter(wbic%in%ctwiWBIC.creel & !(visit.fish.seq.no%in%rm.visit.fish.seq.no)) # reduced interview data
icou.50wd=ccou%>%
  filter(wbic%in%ctwiWBIC.creel & !(visit.fish.seq.no%in%rm.visit.fish.seq.no)) # reduced count data
ifishAg.50wd=cfish%>%
  filter(wbic%in%ctwiWBIC.creel & !(visit.fish.seq.no%in%rm.visit.fish.seq.no)) # reduced aggregate fish data

iharv.50wd=calc_creel_harvest(creel_count_data = icou.50wd,
                              creel_int_data = iint.50wd,
                              creel_fish_data = ifishAg.50wd) # harvest estimates from the reduced data
eff.50wd=calc_creel_effort(creel_count_data = icou.50wd, creel_int_data = iint.50wd)
eff.50wd=eff.50wd%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(annual.eff.acre=total.effort/lake.area)

catch.50wd=iharv.50wd%>%
  select(year,wbic,survey.seq.no, month, daytype, species, species.code, catch, spp.effort, spp.harvest, spp.catch)%>%
  filter(species=='walleye')%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(annual.catch.acre=catch/lake.area)

wd50=ifish.50wd%>%
  group_by(wbic, year,survey.seq.no,survey.begin.date,survey.end.date, mark.found,month,daytype)%>%
  summarise(nfish=sum(fish.count)) # number of fish of each mark typed returned in that creel

wd50=wd50%>% # now throwing out mark type and just taking overall number of marks recapped that month for that survey.
  group_by(wbic, year, survey.seq.no, survey.begin.date, survey.end.date,month,daytype)%>%
  summarise(n.marks.recapped=sum(nfish[!is.na(mark.found)]),
            totalFishExamined=sum(nfish))

# spp.harvest is the column that has the estimate of the harvest for that species in that strata
harvestEstimates.50wd=iharv.50wd%>%
  filter(species.code=="X22")%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest,harvest,total.harvest,total.spp.harvest)%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(harvAcre=spp.harvest/lake.area)

harvestR.50wd=calc_creel_harvest_rates(ifishAg.50wd)%>%
  filter(species.code=="X22")%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest.rate,harvest.rate)

# combining harvest estimates with marked proportion estimate to get total number of marks harvested

markHarvEst.50wd=wd50%>%
  left_join(harvestEstimates.50wd)%>%
  mutate(markHarvEstimate=(n.marks.recapped/totalFishExamined)*spp.harvest)

ang.exp.50wd=markHarvEst.50wd%>%
  left_join(fdat, by=c("wbic"="WBIC","year"="Survey.Year"))%>%
  mutate(exp.rate=markHarvEstimate/nFN.marked)

# Throwing out NA u's and those >1 (NAs from lack of walleye harvest per strata or missing FN marks, >1's is 6 observations where # marks in creel exceeds # marks at large)
ang.exp.50wd=ang.exp.50wd[!is.na(ang.exp.50wd$exp.rate) & ang.exp.50wd$exp.rate<1.0,]

# creating survey-level exploitation rates
ang.exp.50wd=ang.exp.50wd%>%
  group_by(survey.seq.no)%>%
  summarise(meanU=mean(exp.rate,na.rm = T),
            exp.rate.sd=sd(exp.rate,na.rm = T))

###### 50% WEEKENDS ####

rm.fish.seq.no=NA
rm.visit.fish.seq.no=NA
ifish.surv=unique(ifish$survey.seq.no)

set.seed(3)
for(i in 1:length(ifish.surv)){ # survey loop
  full=ifish[ifish$survey.seq.no==ifish.surv[i] & ifish$daytype=="weekend",]
  ms=unique(full$month)
  if(nrow(full)>0){
    for(m in 1:length(ms)){ # month loop
      visits=unique(full$visit.fish.seq.no[full$month==ms[m]])
      visit.rm=visits[c(round(runif(round(0.50*length(visits)),min=1,max = length(visits))))]
      t.rm=full$fish.data.seq.no[full$visit.fish.seq.no%in%visit.rm]# fish seq nos to remove that associated with the visit seq nos selected for removal
      
      rm.visit.fish.seq.no=c(rm.visit.fish.seq.no,visit.rm) # visit fish seq nos to use to reduce the interview, count, and fish data needed to get harvest estimates
      rm.fish.seq.no=c(rm.fish.seq.no,t.rm) # vector of fish to remove from individual fish data
      
    }
  }
}

ifish.50we=ifish[!(ifish$fish.data.seq.no%in%rm.fish.seq.no),] # reduced individual fish data
iint.50we=cint%>%
  filter(wbic%in%ctwiWBIC.creel & !(visit.fish.seq.no%in%rm.visit.fish.seq.no)) # reduced interview data
icou.50we=ccou%>%
  filter(wbic%in%ctwiWBIC.creel & !(visit.fish.seq.no%in%rm.visit.fish.seq.no)) # reduced count data
ifishAg.50we=cfish%>%
  filter(wbic%in%ctwiWBIC.creel & !(visit.fish.seq.no%in%rm.visit.fish.seq.no)) # reduced aggregate fish data

iharv.50we=calc_creel_harvest(creel_count_data = icou.50we,
                              creel_int_data = iint.50we,
                              creel_fish_data = ifishAg.50we) # harvest estimates from the reduced data
eff.50we=calc_creel_effort(creel_count_data = icou.50we, creel_int_data = iint.50we)
eff.50we=eff.50we%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(annual.eff.acre=total.effort/lake.area)

catch.50we=iharv.50we%>%
  select(year,wbic,survey.seq.no, month, daytype, species, species.code, catch, spp.effort, spp.harvest, spp.catch)%>%
  filter(species=='walleye')%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(annual.catch.acre=catch/lake.area)

we50=ifish.50we%>%
  group_by(wbic, year,survey.seq.no,survey.begin.date,survey.end.date, mark.found,month,daytype)%>%
  summarise(nfish=sum(fish.count)) # number of fish of each mark typed returned in that creel

we50=we50%>% # now throwing out mark type and just taking overall number of marks recapped that month for that survey.
  group_by(wbic, year, survey.seq.no, survey.begin.date, survey.end.date,month,daytype)%>%
  summarise(n.marks.recapped=sum(nfish[!is.na(mark.found)]),
            totalFishExamined=sum(nfish))

# spp.harvest is the column that has the estimate of the harvest for that species in that strata
harvestEstimates.50we=iharv.50we%>%
  filter(species.code=="X22")%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest,harvest,total.harvest,total.spp.harvest)%>%
  left_join(lchar[,c(1,15)])%>%
  mutate(harvAcre=spp.harvest/lake.area)

harvestR.50we=calc_creel_harvest_rates(ifishAg.50we)%>%
  filter(species.code=="X22")%>%
  select(year,wbic,survey.seq.no, month, daytype,spp.harvest.rate,harvest.rate)

# combining harvest estimates with marked proportion estimate to get total number of marks harvested

markHarvEst.50we=we50%>%
  left_join(harvestEstimates.50we)%>%
  mutate(markHarvEstimate=(n.marks.recapped/totalFishExamined)*spp.harvest)

ang.exp.50we=markHarvEst.50we%>%
  left_join(fdat, by=c("wbic"="WBIC","year"="Survey.Year"))%>%
  mutate(exp.rate=markHarvEstimate/nFN.marked)

# Throwing out NA u's and those >1 (NAs from lack of walleye harvest per strata or missing FN marks, >1's is 6 observations where # marks in creel exceeds # marks at large)
ang.exp.50we=ang.exp.50we[!is.na(ang.exp.50we$exp.rate) & ang.exp.50we$exp.rate<1.0,]

# creating survey-level exploitation rates
ang.exp.50we=ang.exp.50we%>%
  group_by(survey.seq.no)%>%
  summarise(meanU=mean(exp.rate,na.rm = T),
            exp.rate.sd=sd(exp.rate,na.rm = T))

```



```{r monthlyHarv,fig.width=8,fig.cap="Distribution of harvest rates by month (panel a) for all years of CTWI creel data. Horizontal red line is overall mean harvest rate. Black points are outliers. Panel B is the same information presented on a log-scale. "}


# plot of harvest by month
iharvR$plot.month=month(iharvR$month,label = T)
p1=ggplot(iharvR)+theme_classic()+
  geom_boxplot(aes(y=iharvR$spp.harvest.rate, x=plot.month), fill="grey")+
  geom_hline(yintercept = mean(iharvR$spp.harvest.rate, na.rm = T), color="red")+
  labs(y="Monthly WAE Harvest Rate", x="Month")

# plot of harvest by month - logged
p2=ggplot(iharvR)+theme_classic()+
  geom_boxplot(aes(y=log(iharvR$spp.harvest.rate), x=plot.month), fill="grey")+
  geom_hline(yintercept = log(mean(iharvR$spp.harvest.rate[iharvR$spp.harvest.rate!=0], na.rm = T)), color="red")+ # setting this manually since the 0 exp rates mess up the logging here
  labs(y="Log(Monthly WAE Harvest Rate)", x="Month")

ggarrange(p1,p2,nrow = 1,ncol=2,labels = 'auto')
```

```{r harvPlot, fig.cap="Distribution, on a log scale, of harvest rate for the full data set plus 5 scenarios with reduced data. Actual = full data set, mayAug = May to August creel data only, noWinter = winter creel data removed, wd25 = 25% of weekday creel visits per month removed, wd50 = 50% of weekday creel visits per month removed, we50 = 50% of weekend creel visits per month removed."}

# creating survey-level mean harvest rate for the actual data, estimate of total harvest for that creel year produced from the observed harvest data
harv.surv.a=iharvR%>%
  group_by(survey.seq.no)%>%
  summarise(meanHarvR=mean(spp.harvest.rate,na.rm = T),
            sdHarvR=sd(spp.harvest.rate,na.rm =))
harv.surv.nw=harvestR.nw%>%
  group_by(survey.seq.no)%>%
  summarise(meanHarvR=mean(spp.harvest.rate,na.rm = T),
            sdHarvR=sd(spp.harvest.rate,na.rm = T))
harv.surv.ma=harvestR.ma%>%
  group_by(survey.seq.no)%>%
  summarise(meanHarvR=mean(spp.harvest.rate,na.rm = T),
            sdHarvR=sd(spp.harvest.rate,na.rm = T))
harv.surv.25wd=harvestR.25wd%>%
  group_by(survey.seq.no)%>%
  summarise(meanHarvR=mean(spp.harvest.rate,na.rm = T),
            sdHarvR=sd(spp.harvest.rate,na.rm = T))
harv.surv.50wd=harvestR.50wd%>%
  group_by(survey.seq.no)%>%
  summarise(meanHarvR=mean(spp.harvest.rate,na.rm = T),
            sdHarvR=sd(spp.harvest.rate,na.rm = T))
harv.surv.50we=harvestR.50we%>%
  group_by(survey.seq.no)%>%
  summarise(meanHarvR=mean(spp.harvest.rate,na.rm = T),
            sdHarvR=sd(spp.harvest.rate,na.rm = T))

# combining actual and reduced dataframes into one big one for plotting

ttrExp=rbind(cbind(harv.surv.a,treat=rep("actual",nrow(harv.surv.a))),
             cbind(harv.surv.nw, treat=rep("noWinter",nrow(harv.surv.nw))),
             cbind(harv.surv.ma, treat=rep("mayAug",nrow(harv.surv.ma))),
             cbind(harv.surv.25wd, treat=rep("wd25",nrow(harv.surv.25wd))),
             cbind(harv.surv.50wd, treat=rep("wd50",nrow(harv.surv.50wd))),
             cbind(harv.surv.50we, treat=rep("we50",nrow(harv.surv.50we))))

ggplot(ttrExp)+theme_classic()+
  geom_density(aes(x=log(meanHarvR), fill=treat),alpha=0.2)+
  labs(x="Log(Mean WAE Harvest Rate)", y="Density", fill="Scenario")
# data is non normally distributed, harvest is continuous and greater than 0, consider gamma or lognormal distribution for modeling
```

### Bayesian Model Fitting

Individual likelihoods were constructed for each reduced data set and fit using the functions in the `BayesianTools` package in `R`.
These model fits were checked for convergence using visual inspection of trace plots, posterior distributions, and Gelman-Rubin Diagnostic tests to ensure that models had converged.
All models were fit using Differential Evolution Markov-Chain-Monte-Carlo algorithms run for 10,000 iterations with the first 5,000 discarded as burn-in.

```{r harv-bt-model-fits, echo=T}
# one likelihood to estimate parms for using the 6 different treatments
# creating data frame to model with harvest estimates

#removing 0s since they can't be logged and I'm going to operate under the assumption that if a survey gives a 0 harvest estimate with the full survey, then harvs collecting less data isn't going to change that number. 
zeros.actual=sum(ttrExp$meanHarvR[ttrExp$treat=="actual"]==0)
zeros.nw=sum(ttrExp$meanHarvR[ttrExp$treat=="noWinter"]==0)
zeros.ma=sum(ttrExp$meanHarvR[ttrExp$treat=="mayAug"]==0)
zeros.wd25=sum(ttrExp$meanHarvR[ttrExp$treat=="wd25"]==0)
zeros.wd50=sum(ttrExp$meanHarvR[ttrExp$treat=="wd50"]==0)
zeros.we50=sum(ttrExp$meanHarvR[ttrExp$treat=="we50"]==0)

Ztab=data.frame(Scenario=c("Actual","No Winter","May-August","25% Weekday Reduction","50% Weekday Reduction","50% Weekend Reduction"),
                Zeros=c(zeros.actual,zeros.nw,zeros.ma,zeros.wd25,zeros.wd50,zeros.we50))
kable(Ztab,
      format='latex',
      caption="Table of the number of surveys with harvest estimates equal to 0 for each data scenario.")

# removing 0s here as described in the text.
modDat=ttrExp[ttrExp$meanHarvR!=0 & !is.na(ttrExp$meanHarvR),]

#### ACTUAL ####
harvLL.a=function(param){
  alpha=param[1]
  beta=param[2]

  harvs=rlnorm(nrow(modDat[modDat$treat=="actual",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(modDat$meanHarvR[modDat$treat=="actual"], meanlog = mean(log(harvs)), sdlog = sd(log(harvs)), log = T)
  return(sum(ll))
}

prior=createTruncatedNormalPrior(mean=c(mean(log(modDat$meanHarvR[modDat$treat=="actual"])),sd(log(modDat$meanHarvR[modDat$treat=="actual"]))),
                                 sd=c(1,1),
                                 lower = c(-15,0),
                                 upper = c(15,5))

setup.actual=createBayesianSetup(harvLL.a, prior = prior)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
set.seed(10)
harv.actual=runMCMC(bayesianSetup = setup.actual, sampler = "DEzs", settings = settings) # takes about 10 seconds

#### NW ####
harvLL.nw=function(param){
  alpha=param[1]
  beta=param[2]
  
  harvs=rlnorm(nrow(modDat[modDat$treat=="noWinter",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(modDat$meanHarvR[modDat$treat=="noWinter"], meanlog = mean(log(harvs)), sdlog = sd(log(harvs)), log = T)
  return(sum(ll))
}

prior=createTruncatedNormalPrior(mean=c(mean(log(modDat$meanHarvR[modDat$treat=="noWinter"])),sd(log(modDat$meanHarvR[modDat$treat=="noWinter"]))), 
                                 sd=c(1,1),
                                 lower = c(-15,0),
                                 upper = c(15,5))

setup.nw=createBayesianSetup(harvLL.nw, prior = prior)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
set.seed(10)
harv.nw=runMCMC(bayesianSetup = setup.nw, sampler = "DEzs", settings = settings)

#### MA ####

harvLL.ma=function(param){
  alpha=param[1]
  beta=param[2]
  
  harvs=rlnorm(nrow(modDat[modDat$treat=="mayAug",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(modDat$meanHarvR[modDat$treat=="mayAug"], meanlog = mean(log(harvs)), sdlog = sd(log(harvs)), log = T)
  return(sum(ll))
}

prior=createTruncatedNormalPrior(mean=c(mean(log(modDat$meanHarvR[modDat$treat=="mayAug"])),sd(log(modDat$meanHarvR[modDat$treat=="mayAug"]))), 
                                 sd=c(1,1),
                                 lower = c(-15,0),
                                 upper = c(15,5))

setup.ma=createBayesianSetup(harvLL.ma, prior = prior)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
set.seed(10)
harv.ma=runMCMC(bayesianSetup = setup.ma, sampler = "DEzs", settings = settings)

#### WD25 ####

harvLL.wd25=function(param){
  alpha=param[1]
  beta=param[2]
  
  harvs=rlnorm(nrow(modDat[modDat$treat=="wd25",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(modDat$meanHarvR[modDat$treat=="wd25"], meanlog = mean(log(harvs)), sdlog = sd(log(harvs)), log = T)
  return(sum(ll))
}

prior=createTruncatedNormalPrior(mean=c(mean(log(modDat$meanHarvR[modDat$treat=="wd25"])),sd(log(modDat$meanHarvR[modDat$treat=="wd25"]))), 
                                 sd=c(1,1),
                                 lower = c(-15,0),
                                 upper = c(15,5))

setup.wd25=createBayesianSetup(harvLL.wd25, prior = prior)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
set.seed(10)
harv.25wd=runMCMC(bayesianSetup = setup.wd25, sampler = "DEzs", settings = settings)

#### WD50 ####
harvLL.wd50=function(param){
  alpha=param[1]
  beta=param[2]
  
  harvs=rlnorm(nrow(modDat[modDat$treat=="wd50",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(modDat$meanHarvR[modDat$treat=="wd50"], meanlog = mean(log(harvs)), sdlog = sd(log(harvs)), log = T)
  return(sum(ll))
}

prior=createTruncatedNormalPrior(mean=c(mean(log(modDat$meanHarvR[modDat$treat=="wd50"])),sd(log(modDat$meanHarvR[modDat$treat=="wd50"]))), 
                                 sd=c(1,1),
                                 lower = c(-15,0),
                                 upper = c(15,5))

setup.wd50=createBayesianSetup(harvLL.wd50, prior = prior)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
set.seed(10)
harv.50wd=runMCMC(bayesianSetup = setup.wd50, sampler = "DEzs", settings = settings)

#### WE50 ####
harvLL.we50=function(param){
  alpha=param[1]
  beta=param[2]
  
  harvs=rlnorm(nrow(modDat[modDat$treat=="we50",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(modDat$meanHarvR[modDat$treat=="we50"], meanlog = mean(log(harvs)), sdlog = sd(log(harvs)), log = T)
  return(sum(ll))
}

prior=createTruncatedNormalPrior(mean=c(mean(log(modDat$meanHarvR[modDat$treat=="we50"])),sd(log(modDat$meanHarvR[modDat$treat=="we50"]))), 
                                 sd=c(1,1),
                                 lower = c(-15,0),
                                 upper = c(15,5))

setup.we50=createBayesianSetup(harvLL.we50, prior = prior)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
set.seed(10)
harv.50we=runMCMC(bayesianSetup = setup.we50, sampler = "DEzs", settings = settings)

```

### Model Checking


```{r harvDataReproduction, fig.width=8, fig.height=6 ,fig.cap="Distributions of the observed and model predicted data for each scenario. Model predicted data comes from lognormal distribution paramterized using the median parameter estimate of the model posterior."}

# looking to see if parm estimates produce data that visually at least looks like the observed data for that scenario
pars.a=getSample(harv.actual)
pars.nw=getSample(harv.nw)
pars.ma=getSample(harv.ma)
pars.wd25=getSample(harv.25wd)
pars.wd50=getSample(harv.50wd)
pars.we50=getSample(harv.50we)

set.seed(3)
aComp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="actual"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="actual"]),
                                                                         meanlog = median(pars.a[,1]),
                                                                         sdlog = median(pars.a[,2]))),
                    treat=c(rep("observed",length(modDat$meanHarvR[modDat$treat=="actual"])),rep("pred",length(modDat$meanHarvR[modDat$treat=="actual"]))))
set.seed(3)
nwComp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="noWinter"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="noWinter"]),
                                                                    meanlog = median(pars.nw[,1]),
                                                                    sdlog = median(pars.nw[,2]))),
                 treat=c(rep("observed",length(modDat$meanHarvR[modDat$treat=="noWinter"])),rep("pred",length(modDat$meanHarvR[modDat$treat=="noWinter"]))))
set.seed(3)
maComp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="mayAug"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="mayAug"]),
                                                                    meanlog = median(pars.ma[,1]),
                                                                    sdlog = median(pars.ma[,2]))),
                 treat=c(rep("observed",length(modDat$meanHarvR[modDat$treat=="mayAug"])),rep("pred",length(modDat$meanHarvR[modDat$treat=="mayAug"]))))
set.seed(3)
wd25Comp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="wd25"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="wd25"]),
                                                                    meanlog = median(pars.wd25[,1]),
                                                                    sdlog = median(pars.wd25[,2]))),
                 treat=c(rep("observed",length(modDat$meanHarvR[modDat$treat=="wd25"])),rep("pred",length(modDat$meanHarvR[modDat$treat=="wd25"]))))
set.seed(3)
wd50Comp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="wd50"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="wd50"]),
                                                                    meanlog = median(pars.wd50[,1]),
                                                                    sdlog = median(pars.wd50[,2]))),
                 treat=c(rep("observed",length(modDat$meanHarvR[modDat$treat=="wd50"])),rep("pred",length(modDat$meanHarvR[modDat$treat=="wd50"]))))
set.seed(3)
we50Comp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="we50"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="we50"]),
                                                                    meanlog = median(pars.we50[,1]),
                                                                    sdlog = median(pars.we50[,2]))),
                 treat=c(rep("observed",length(modDat$meanHarvR[modDat$treat=="we50"])),rep("pred",length(modDat$meanHarvR[modDat$treat=="we50"]))))

a.p=ggplot(aComp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "Actual", fill=element_blank())
nw.p=ggplot(nwComp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "No Winter Data \n(April - October)", fill=element_blank())
ma.p=ggplot(maComp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "May - August \nData Only", fill=element_blank())
wd25.p=ggplot(wd25Comp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "25% of Weekday \nCreel Visits/Month \nRemoved", fill=element_blank())
wd50.p=ggplot(wd25Comp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "50% of Weekday \nCreel Visits/Month \nRemoved", fill=element_blank())
we50.p=ggplot(we50Comp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "50% of Weekdend \nCreel Visits/Month \nRemoved", fill=element_blank())

ggarrange(a.p,nw.p,ma.p,wd25.p,wd50.p,we50.p, common.legend = T)
```

```{r harv-gelman-rubin-diagnostics, echo=T, eval=FALSE}

gelmanDiagnostics(harv.actual) # converged 
gelmanDiagnostics(harv.nw) # converged
gelmanDiagnostics(harv.ma) # converged
gelmanDiagnostics(harv.25wd) # converged
gelmanDiagnostics(harv.50wd) # converged
gelmanDiagnostics(harv.50we) # converged

```

```{r harvbpValue}

# dataframe to hold output
bpValues=data.frame(scenario=c("Actual", "No Winter", "May-August","25% weeday removal","50% weekday removal","50% weekend removal"),
                    sd.pval=NA)
#### Pb ACTUAL ####
pval.actual=data.frame(alpha=rep(NA,nrow(pars.a)),
                       beta=NA,
                       sd=NA)
set.seed(10)
for(i in 1:nrow(pars.a)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="actual"]),
                 meanlog = pars.a[i,1],
                 sdlog = pars.a[i,2])
  pval.actual$alpha[i]=pars.a[i,1]
  pval.actual$beta[i]=pars.a[i,2]
  pval.actual$sd[i]=sd(log(tempdat))
}

# now calculate the number of times sd exceeds that of the real data

pval.actual$sdExceed=0

actual.sd=sd(log(modDat$meanHarvR[modDat$treat=="actual"]))

pval.actual$sdExceed[pval.actual$sd>actual.sd]=1

bpValues$sd.pval[1]=sum(pval.actual$sdExceed==1)/nrow(pval.actual) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

#### Pb NO WINTER ####
pval.noWinter=data.frame(alpha=rep(NA,nrow(pars.nw)),
                       beta=NA,
                       sd=NA)
set.seed(10)
for(i in 1:nrow(pars.nw)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="noWinter"]),
                 meanlog = pars.nw[i,1],
                 sdlog = pars.nw[i,2])
  pval.noWinter$alpha[i]=pars.nw[i,1]
  pval.noWinter$beta[i]=pars.nw[i,2]
  pval.noWinter$sd[i]=sd(log(tempdat))
}

# now calculate the number of times the sd exceeds that of the real data

pval.noWinter$sdExceed=0

noWinter.sd=sd(log(modDat$meanHarvR[modDat$treat=="noWinter"]))

pval.noWinter$sdExceed[pval.noWinter$sd>noWinter.sd]=1

bpValues$sd.pval[2]=sum(pval.noWinter$sdExceed==1)/nrow(pval.noWinter) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

#### Pb MAYAUGUST ####
pval.mayAug=data.frame(alpha=rep(NA,nrow(pars.ma)),
                       beta=NA,
                       sd=NA)
set.seed(10)
for(i in 1:nrow(pars.ma)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="mayAug"]),
                 meanlog = pars.ma[i,1],
                 sdlog = pars.ma[i,2])
  pval.mayAug$alpha[i]=pars.ma[i,1]
  pval.mayAug$beta[i]=pars.ma[i,2]
  pval.mayAug$sd[i]=sd(log(tempdat))
}

# now calculate the number of times the sd exceeds that of the real data

pval.mayAug$sdExceed=0

mayAug.sd=sd(log(modDat$meanHarvR[modDat$treat=="mayAug"]))

pval.mayAug$sdExceed[pval.mayAug$sd>mayAug.sd]=1

bpValues$sd.pval[3]=sum(pval.mayAug$sdExceed==1)/nrow(pval.mayAug) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

#### Pb WD25 ####
pval.wd25=data.frame(alpha=rep(NA,nrow(pars.wd25)),
                       beta=NA,
                       sd=NA)
set.seed(10)
for(i in 1:nrow(pars.wd25)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="wd25"]),
                 meanlog = pars.wd25[i,1],
                 sdlog = pars.wd25[i,2])
  pval.wd25$alpha[i]=pars.wd25[i,1]
  pval.wd25$beta[i]=pars.wd25[i,2]
  pval.wd25$sd[i]=sd(log(tempdat))
}

# now calculate the number of times the sd exceeds that of the real data

pval.wd25$sdExceed=0

wd25.sd=sd(log(modDat$meanHarvR[modDat$treat=="wd25"]))

pval.wd25$sdExceed[pval.wd25$sd>wd25.sd]=1

bpValues$sd.pval[4]=sum(pval.wd25$sdExceed==1)/nrow(pval.wd25) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

#### Pb WD50 ####
pval.wd50=data.frame(alpha=rep(NA,nrow(pars.wd50)),
                       beta=NA,
                       sd=NA)
set.seed(10)
for(i in 1:nrow(pars.wd50)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="wd50"]),
                 meanlog = pars.wd50[i,1],
                 sdlog = pars.wd50[i,2])
  pval.wd50$alpha[i]=pars.wd50[i,1]
  pval.wd50$beta[i]=pars.wd50[i,2]
  pval.wd50$sd[i]=sd(log(tempdat))
}

# now calculate the number of times the sd exceeds that of the real data

pval.wd50$sdExceed=0

wd50.sd=sd(log(modDat$meanHarvR[modDat$treat=="wd50"]))

pval.wd50$sdExceed[pval.wd50$sd>wd50.sd]=1

bpValues$sd.pval[5]=sum(pval.wd50$sdExceed==1)/nrow(pval.wd50) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

#### Pb WE50 ####
pval.we50=data.frame(alpha=rep(NA,nrow(pars.we50)),
                       beta=NA,
                       sd=NA)
set.seed(10)
for(i in 1:nrow(pars.we50)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="we50"]),
                 meanlog = pars.we50[i,1],
                 sdlog = pars.we50[i,2])
  pval.we50$alpha[i]=pars.we50[i,1]
  pval.we50$beta[i]=pars.we50[i,2]
  pval.we50$sd[i]=sd(log(tempdat))
}

# now calculate the number of times the sd exceeds that of the real data

pval.we50$sdExceed=0

we50.sd=sd(log(modDat$meanHarvR[modDat$treat=="we50"]))

pval.we50$sdExceed[pval.we50$sd>we50.sd]=1

bpValues$sd.pval[6]=sum(pval.we50$sdExceed==1)/nrow(pval.we50) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

kable(bpValues, digits = 3, col.names = c("Scenario", "SD p-value"), align="lcc",caption = "Bayesian p-values for a lognormal distributed model. Each p-value describes whether or not there is a significant difference between data generated by the fitted model and the data for that scenario. Standard devidation (SD) was the metric assessed here.")
```

### Inference

Having established that the models are fitting the data well, some inference can be gained as to whether or not the reductions in creel effort proposed here would result in a meaningful change in the total harvest rate estimated from the reduced data.

```{r harvDataComparisonToActual, fig.width=8,fig.height=6,fig.cap="Comparison of the distribution of actual harvest rates calculated from creel data and harvest rate calculated from simulated creel data for each data reduction scenario and the resulting median parameter values for their respective model fits."}

aComp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="actual"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="actual"]),
                                                                    meanlog = median(pars.a[,1]),
                                                                    sdlog = median(pars.a[,2]))),
                 treat=c(rep("actual",length(modDat$meanHarvR[modDat$treat=="actual"])),rep("model",length(modDat$meanHarvR[modDat$treat=="actual"]))))

nwComp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="actual"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="actual"]),
                                                                       meanlog = median(pars.nw[,1]),
                                                                       sdlog = median(pars.nw[,2]))),
                  treat=c(rep("actual",length(modDat$meanHarvR[modDat$treat=="actual"])),rep("model",length(modDat$meanHarvR[modDat$treat=="actual"]))))

maComp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="actual"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="actual"]),
                                                                     meanlog = median(pars.ma[,1]),
                                                                     sdlog = median(pars.ma[,2]))),
                  treat=c(rep("actual",length(modDat$meanHarvR[modDat$treat=="actual"])),rep("model",length(modDat$meanHarvR[modDat$treat=="actual"]))))

wd25Comp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="actual"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="actual"]),
                                                                     meanlog = median(pars.wd25[,1]),
                                                                     sdlog = median(pars.wd25[,2]))),
                    treat=c(rep("actual",length(modDat$meanHarvR[modDat$treat=="actual"])),rep("model",length(modDat$meanHarvR[modDat$treat=="actual"]))))

wd50Comp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="actual"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="actual"]),
                                                                     meanlog = median(pars.wd50[,1]),
                                                                     sdlog = median(pars.wd50[,2]))),
                    treat=c(rep("actual",length(modDat$meanHarvR[modDat$treat=="actual"])),rep("model",length(modDat$meanHarvR[modDat$treat=="actual"]))))

we50Comp=data.frame(harvR=c(modDat$meanHarvR[modDat$treat=="actual"],rlnorm(n=length(modDat$meanHarvR[modDat$treat=="actual"]),
                                                                     meanlog = median(pars.we50[,1]),
                                                                     sdlog = median(pars.we50[,2]))),
                    treat=c(rep("actual",length(modDat$meanHarvR[modDat$treat=="actual"])),rep("model",length(modDat$meanHarvR[modDat$treat=="actual"]))))

a.p=ggplot(aComp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "Actual", fill=element_blank())
nw.p=ggplot(nwComp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "No Winter Data \n(April - October)", fill=element_blank())
ma.p=ggplot(maComp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "May - August \nData Only", fill=element_blank())
wd25.p=ggplot(wd25Comp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "25% of Weekday \nCreel Visits/Month \nRemoved", fill=element_blank())
wd50.p=ggplot(wd25Comp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "50% of Weekday \nCreel Visits/Month \nRemoved", fill=element_blank())
we50.p=ggplot(we50Comp)+theme_classic()+
  geom_density(aes(x=log(harvR),fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Log(Mean WAE Harvest Rate)", y="Density", title = "50% of Weekdend \nCreel Visits/Month \nRemoved", fill=element_blank())

ggarrange(a.p,nw.p,ma.p,wd25.p,wd50.p,we50.p, common.legend = T)
```

A Bayesian p-value can be employed here too.
A test between the sd of the actual data and the sd of the reduced data can describe whether the model of the reduced data is able to approximate the actual data or not.
Successful approximations are p-values close to 0.5 with reduced fit as values increase or decrease beyond 0.5.
Generally, the rule of thumb is that p-values $<0.10$ or $>0.90$ signal unacceptable fits.
However, these cutoffs can be set based on the objectives of the study at hand and the values of the decision makers.



```{r bpValueComparison, warning=F, message=F}
## p-value tests comparing the scenarios to actual to show that some scenarios approximate the actual quite well.

bpval.comp=data.frame(scenario=c("Actual", "No Winter", "May-August","25% weeday removal","50% weekday removal","50% weekend removal"),
                      sd.pval=NA)
pval.actual$sdComp=0
pval.actual$sdComp[pval.actual$sd>actual.sd]=1

pval.noWinter$sdComp=0
pval.noWinter$sdComp[pval.noWinter$sd>actual.sd]=1

pval.mayAug$sdComp=0
pval.mayAug$sdComp[pval.mayAug$sd>actual.sd]=1

pval.wd25$sdComp=0
pval.wd25$sdComp[pval.wd25$sd>actual.sd]=1

pval.wd50$sdComp=0
pval.wd50$sdComp[pval.wd50$sd>actual.sd]=1

pval.we50$sdComp=0
pval.we50$sdComp[pval.we50$sd>actual.sd]=1


bpval.comp$sd.pval=c(sum(pval.actual$sdComp)/nrow(pval.actual),
                           sum(pval.noWinter$sdComp)/nrow(pval.noWinter),
                           sum(pval.mayAug$sdComp)/nrow(pval.mayAug),
                           sum(pval.wd25$sdComp)/nrow(pval.wd25),
                           sum(pval.wd50$sdComp)/nrow(pval.wd50),
                           sum(pval.we50$sdComp)/nrow(pval.we50))

kable(bpval.comp, digits = 3, col.names = c("Scenario", "SD p-value"), align="lcc",caption = "Bayesian p-values for the comparison between the actual data and the scenario-specific model fit to a lognormal distribution. This test is asking whether the scenario-specific model can approximate the actual data. When true this signals no effect of the data reduction for that scenario on the resulting harvest rate estimate. Standard devidation (SD) was the metric assessed here.")
```

### Year Specific Analyses

Instead of modeling the full population of harvest rates as was done above, here I have done the model fitting on individual creel-years so that I can see how the removal of data for a creel-year (which is a comparatively large amount of missing data since only 16-20 lakes are creeled each year) impacts the harvest rate that year.
This analysis will follow the same routine as the whole-population analyses above but only consider one creel-year at a time.


```{r yearAnalysisExampleLakes, fig.width=8,fig.cap="Comparison of the harvest rate from the reduced data for a handful of creel surveys. Most data reductions do not result in harvest rate that are much different from the actual data (most are within 1 SD of the harvest estimate for that treatment). Colors represent data reduction scenarios, points are mean harevst rate across all months of that creel survey and vertical lines represent 1 standard deviation."}
ttrExp=ttrExp%>%
  left_join(cserv[,2:5])%>%
  left_join(lchar[,c(1,3,15)])

pdat=ttrExp[ttrExp$year%in%c(2015,2019,2022),]
ggplot(pdat)+theme_classic()+
  geom_pointrange(aes(x=paste(year,waterbody.name,sep = "_"), 
                      y=meanHarvR, ymin=meanHarvR-sdHarvR, 
                      ymax=meanHarvR+sdHarvR, color=treat), 
                  position = position_dodge(width = 1))+
  theme(axis.text.x = element_text(angle=45,hjust=1), legend.position = c(.75,.75))+
  labs(x="Lake-Year",y="Mean WAE Harvest Rate (+/- 1 SD)",color="Scenario")+
  scale_color_viridis_d()+coord_cartesian(ylim = c(0,NA))

```

```{r year-analysis-diff-actual, fig.cap="Boxplot of the difference between the actual harvest rate and the data-reduced harvest rate for each creel survey in the creel dataset. Most of the data exhibits differences very near 0 except the seasonal reductions, dots represent outliers (x<|> x's percentile-1.5*interquartile range)"}

# metrics comparing differences in individual lake estimates

trLake=ttrExp
trLake$a.diff=NA
trLake$a.exceed=NA

for(i in 1:nrow(trLake)){
  trLake$a.diff[i]=trLake$meanHarvR[trLake$treat=="actual" & trLake$survey.seq.no==trLake$survey.seq.no[i]]-trLake$meanHarvR[i]
  trLake$a.exceed[i]=trLake$a.diff[i]>=trLake$sdHarvR[trLake$treat=="actual" & trLake$survey.seq.no==trLake$survey.seq.no[i]]
}

ggplot(trLake)+theme_classic()+
   geom_boxplot(aes(y=a.diff, x=treat))+
  labs(x="Data Reduction Scenario",y="Difference from actual")
```

```{r year-analysis-exceedenceTable, eval=F}

exceedSummary=trLake%>%
  group_by(treat)%>%
  summarise(nTrue=sum(a.exceed,na.rm = T),
            nFalse=sum(a.exceed==F,na.rm = T),
            nNA=sum(is.na(a.exceed)))
kable(exceedSummary,
      col.names = c("Data Reduction Scenario","Harvet estimates exceeding actual","Harvest estimates not exceeding actual","NAs"),
      caption = "Counts of the number of times the harvest rate for the reduced data scenario differs from the actual harvest rate for that treatment by more than 1 standard deviation. NAs ocurred where there was no variance in the harvest rate estimate and thus no standard deviation could be calculated for the actual data.")

```

```{r yearAnalysisRecCode, fig.cap="Comparison of the magnitutde of the difference between the mean actual harvest rate and the estimate derived from each reduced data set. Vertical lines represent 1 standard deviation. Walleye recruitment codes are along the x axis. Note the y-axis scale, many of the differences are small. Walleye recruitment codes describe the source of young-of-year walleye in the system and are defined as follows: c=combined equal shares stocked and natural source, cnr=combined mainly from natural, cst=combined mainly from stocking, nr=natural reproduction only, ost=stocking only with few/no adult walleye present,, st=stocked only, NA=not evaluated yet."}

# walleye recruitment class from lchar
# walleye recruitment class from lchar
# effort from ceff pooled to survey level
lEff=calc_creel_effort(creel_count_data = ccou,
                       creel_int_data = cint,
                       grouping = c("wbic","survey.seq.no","month","daytype"))
# current lake classes
lc=read.csv('lake class predictions.csv')# these come from Rypel et al. 2019 https://doi.org/10.1002/fsh.10228
lc$LakeClass=gsub(" ","-",lc$LakeClass)

chars=lEff%>%
  group_by(wbic,survey.seq.no)%>%
  summarise(effort_hrs=sum(total.effort),
            effort_hrs.sd=sd(total.effort,na.rm=T))%>%
  left_join(lchar[,c(1,3,15,35)])%>%
  mutate(effortHrs.hectare=effort_hrs/(lake.area/2.47),
         effortHrs.hectare.sd=sd(effort_hrs/(lake.area/2.47),na.rm=T))%>%
  left_join(lc[,c(1,10)],by=c('wbic'='WBIC'))
trLake_chars=trLake%>%
  left_join(lc[,c(1,10)],by=c('wbic'='WBIC'))%>%
  left_join(lchar[,c(1,35)])%>%
  left_join(chars)

stClass=trLake_chars%>%
  group_by(wae.code,treat)%>%
  summarise(meanDiff=mean(a.diff,na.rm = T),
            sdDiff=sd(a.diff,na.rm=T))
ggplot(stClass)+theme_classic()+
  geom_pointrange(aes(x=wae.code,y=meanDiff,ymin=meanDiff-sdDiff,ymax=meanDiff+sdDiff,color=treat),position = position_dodge(width = 1))+
  scale_color_viridis_d()+theme(legend.position = c(0.7,0.8))+
  labs(x="WAE Recruitment Code",y="Mean Difference from Actual Harvest Rate",color="Scenario")


```
```{r yearAnalysisLakeClass,fig.cap="Comparison of the magnitude of the difference between the mean actual harvest rate and the estimate derived from each reduced data set. Vertical lines represent 1 standard deviation. Note the y-axis scale, many of the differences are very small. Lake classes are according to Rypel et al. 2019, https://doi.org/10.1002/fsh.10228 ."}
lcSum=trLake_chars%>%
  group_by(LakeClass,treat)%>%
  summarise(meanDiff=mean(a.diff,na.rm = T),
            sdDiff=sd(a.diff,na.rm=T))
ggplot(lcSum)+theme_classic()+
  geom_pointrange(aes(x=LakeClass,y=meanDiff, ymin=meanDiff-sdDiff,ymax=meanDiff+sdDiff,color=treat),position = position_dodge(width = 1))+
  scale_color_viridis_d()+
  theme(axis.text.x = element_text(angle=45,hjust = 1), legend.position = c(0.8,0.75))+
  labs(x="Lake Class",y="Mean Difference from Actual Harvest Rate", color="Scenario")
```

```{r yearAnalysisLakeArea, fig.cap="Comparison of the magnitude of the difference between the mean actual harvest rate and the estimate derived from each reduced data set. Note the y-axis scale, many of the differences are small."}
ggplot(trLake_chars)+theme_classic()+
  geom_point(aes(y=a.diff,x=log(lake.area/2.47),color=treat))+
  scale_color_viridis_d()+theme(legend.position = c(0.8,0.4))+
  labs(x="Log(Lake Area in Hectares)", y="Difference from Actual harvest rate",color="Scenario")
```

```{r yearAnalysisEffDens, fig.cap="Comparison of the magnitude of the difference between the mean actual harvest rate and the estimate derived from each reduced data set. Note the y-axis scale, many of the differences are small."}
ggplot(trLake_chars)+theme_classic()+
  geom_point(aes(y=a.diff,x=log(effortHrs.hectare),color=treat))+
  scale_color_viridis_d()+theme(legend.position = c(0.8,0.4))+
  labs(x="Log(Effort/Hectare)",y="Difference from Actual harvest rate",color="Scenario")
```

```{r year-loop, eval=F, echo=T}
# MODELING EFFECTS OF DATA REDUCTION ON INDIVIDUAL YEARS

# likelihoods to fit
harvLL.we50=function(param){
  alpha=param[1]
  beta=param[2]
  
  hvs=rlnorm(nrow(tdat[tdat$treat=="we50",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(tdat$meanHarvR[tdat$treat=="we50"], meanlog = mean(log(hvs)), sdlog = sd(log(hvs)), log = T)
  return(sum(ll))
}
harvLL.wd50=function(param){
  alpha=param[1]
  beta=param[2]
  
  hvs=rlnorm(nrow(tdat[tdat$treat=="wd50",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(tdat$meanHarvR[tdat$treat=="wd50"], meanlog = mean(log(hvs)), sdlog = sd(log(hvs)), log = T)
  return(sum(ll))
}
harvLL.wd25=function(param){
  alpha=param[1]
  beta=param[2]
  
  hvs=rlnorm(nrow(tdat[tdat$treat=="wd25",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(tdat$meanHarvR[tdat$treat=="wd25"], meanlog = mean(log(hvs)), sdlog = sd(log(hvs)), log = T)
  return(sum(ll))
}
harvLL.ma=function(param){
  alpha=param[1]
  beta=param[2]
  
  hvs=rlnorm(nrow(tdat[tdat$treat=="mayAug",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(tdat$meanHarvR[tdat$treat=="mayAug"], meanlog = mean(log(hvs)), sdlog = sd(log(hvs)), log = T)
  return(sum(ll))
}
harvLL.nw=function(param){
  alpha=param[1]
  beta=param[2]
  
  hvs=rlnorm(nrow(tdat[tdat$treat=="noWinter",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(tdat$meanHarvR[tdat$treat=="noWinter"], meanlog = mean(log(hvs)), sdlog = sd(log(hvs)), log = T)
  return(sum(ll))
}
harvLL.a=function(param){
  alpha=param[1]
  beta=param[2]

  hvs=rlnorm(nrow(tdat[tdat$treat=="actual",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(tdat$meanHarvR[tdat$treat=="actual"], meanlog = mean(log(hvs)), sdlog = sd(log(hvs)), log = T)
  return(sum(ll))
}

# removing 0s and NAs
ttrExp=ttrExp[ttrExp$meanHarvR!=0 & !is.na(ttrExp$meanHarvR),]
loopY=sort(unique(ttrExp$year))

bpval.comp.y=data.frame(year=NA,
                      scenario=NA,
                      sd.pval=NA)

bpval.self.y=data.frame(year=NA,
                        scenario=NA,
                        sd.pval=NA)
grMetrics.y=data.frame(year=NA,
                       scenario=NA,
                       gr.prsf=NA,
                       gr.par1=NA,
                       gr.par2=NA)

for(y in 1:length(loopY)){
  #first get to year-specific data
  tdat=ttrExp[ttrExp$year==loopY[y],]
  # Bayesian Model Fitting
  #ACTUAL

  prior=createTruncatedNormalPrior(mean=c(mean(log(ttrExp$meanHarvR[ttrExp$treat=="actual"])),sd(log(ttrExp$meanHarvR[ttrExp$treat=="actual"]))),
                                   sd=c(1,1),
                                   lower = c(-15,0),
                                   upper = c(15,5))

  setup.a=createBayesianSetup(harvLL.a, prior = prior)

  settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
  set.seed(10)
  t.a=runMCMC(bayesianSetup = setup.a, sampler = "DEzs", settings = settings)
  #NW

  prior=createTruncatedNormalPrior(mean=c(mean(log(ttrExp$meanHarvR[ttrExp$treat=="noWinter"])),sd(log(ttrExp$meanHarvR[ttrExp$treat=="noWinter"]))),
                                   sd=c(1,1),
                                   lower = c(-15,0),
                                   upper = c(15,5))

  setup.nw=createBayesianSetup(harvLL.nw, prior = prior)

  settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
  set.seed(10)
  t.nw=runMCMC(bayesianSetup = setup.nw, sampler = "DEzs", settings = settings)
  #MA

  prior=createTruncatedNormalPrior(mean=c(mean(log(ttrExp$meanHarvR[ttrExp$treat=="mayAug"])),sd(log(ttrExp$meanHarvR[ttrExp$treat=="mayAug"]))),
                                   sd=c(1,1),
                                   lower = c(-15,0),
                                   upper = c(15,5))

  setup.ma=createBayesianSetup(harvLL.ma, prior = prior)

  settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
  set.seed(10)
  t.ma=runMCMC(bayesianSetup = setup.ma, sampler = "DEzs", settings = settings)
  #WD.25

  prior=createTruncatedNormalPrior(mean=c(mean(log(ttrExp$meanHarvR[ttrExp$treat=="wd25"])),sd(log(ttrExp$meanHarvR[ttrExp$treat=="wd25"]))),
                                   sd=c(1,1),
                                   lower = c(-15,0),
                                   upper = c(15,5))

  setup.wd25=createBayesianSetup(harvLL.wd25, prior = prior)

  settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
  set.seed(10)
  t.25wd=runMCMC(bayesianSetup = setup.wd25, sampler = "DEzs", settings = settings)
  #WD.50

  prior=createTruncatedNormalPrior(mean=c(mean(log(ttrExp$meanHarvR[ttrExp$treat=="wd50"])),sd(log(ttrExp$meanHarvR[ttrExp$treat=="wd50"]))),
                                   sd=c(1,1),
                                   lower = c(-15,0),
                                   upper = c(15,5))

  setup.wd50=createBayesianSetup(harvLL.wd50, prior = prior)

  settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
  set.seed(10)
  t.50wd=runMCMC(bayesianSetup = setup.wd50, sampler = "DEzs", settings = settings)

  #WE.50

  prior=createTruncatedNormalPrior(mean=c(mean(log(ttrExp$meanHarvR[ttrExp$treat=="we50"])),sd(log(ttrExp$meanHarvR[ttrExp$treat=="we50"]))),
                                   sd=c(1,1),
                                   lower = c(-15,0),
                                   upper = c(15,5))

  setup.we50=createBayesianSetup(harvLL.we50, prior = prior)

  settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
  set.seed(10)
  t.50we=runMCMC(bayesianSetup = setup.we50, sampler = "DEzs", settings = settings)
  
  ## GR Diagnostics to record for each year
  
  gr.a=gelmanDiagnostics(t.a)
  gr.ma=gelmanDiagnostics(t.ma)
  gr.nw=gelmanDiagnostics(t.nw)
  gr.wd25=gelmanDiagnostics(t.25wd)
  gr.wd50=gelmanDiagnostics(t.50wd)
  gr.we50=gelmanDiagnostics(t.50we)
  
    ## BAYESIAN P-VALUE CALCS to record for each year

  pars.a=getSample(t.a)
  pars.nw=getSample(t.nw)
  pars.ma=getSample(t.ma)
  pars.wd25=getSample(t.25wd)
  pars.wd50=getSample(t.50wd)
  pars.we50=getSample(t.50we)

  #### Pb ACTUAL ####
  pval.actual=data.frame(alpha=rep(NA,nrow(pars.a)),
                         beta=NA,
                         sd=NA)
  set.seed(10)
  for(i in 1:nrow(pars.a)){
    tempdat=rlnorm(n=length(tdat$meanHarvR[tdat$treat=="actual"]),
                   meanlog = pars.a[i,1],
                   sdlog = pars.a[i,2])
    pval.actual$alpha[i]=pars.a[i,1]
    pval.actual$beta[i]=pars.a[i,2]
    pval.actual$sd[i]=sd(log(tempdat))
  }

  # now calculate the number of times the sd exceeds that of the real data

  pval.actual$sdExceed=0

  actual.sd=sd(log(tdat$meanHarvR[tdat$treat=="actual"]))

  pval.actual$sdExceed[pval.actual$sd>actual.sd]=1

  #### Pb NO WINTER ####
  pval.noWinter=data.frame(alpha=rep(NA,nrow(pars.nw)),
                           beta=NA,
                           sd=NA)
  set.seed(10)
  for(i in 1:nrow(pars.nw)){
    tempdat=rlnorm(n=length(tdat$meanHarvR[tdat$treat=="noWinter"]),
                   meanlog = pars.nw[i,1],
                   sdlog = pars.nw[i,2])
    pval.noWinter$alpha[i]=pars.nw[i,1]
    pval.noWinter$beta[i]=pars.nw[i,2]
    pval.noWinter$sd[i]=sd(log(tempdat))
  }

  # now calculate the number of times the sd exceeds that of the real data

  pval.noWinter$sdExceed=0

  noWinter.sd=sd(log(tdat$meanHarvR[tdat$treat=="noWinter"]))

  pval.noWinter$sdExceed[pval.noWinter$sd>noWinter.sd]=1

  #### Pb MAYAUGUST ####
  pval.mayAug=data.frame(alpha=rep(NA,nrow(pars.ma)),
                         beta=NA,
                         sd=NA)
  set.seed(10)
  for(i in 1:nrow(pars.ma)){
    tempdat=rlnorm(n=length(tdat$meanHarvR[tdat$treat=="mayAug"]),
                   meanlog = pars.ma[i,1],
                   sdlog = pars.ma[i,2])
    pval.mayAug$alpha[i]=pars.ma[i,1]
    pval.mayAug$beta[i]=pars.ma[i,2]
    pval.mayAug$sd[i]=sd(log(tempdat))
  }

  # now calculate the number of times the sd exceeds that of the real data

  pval.mayAug$sdExceed=0

  mayAug.sd=sd(log(tdat$meanHarvR[tdat$treat=="mayAug"]))

  pval.mayAug$sdExceed[pval.mayAug$sd>mayAug.sd]=1

  #### Pb WD25 ####
  pval.wd25=data.frame(alpha=rep(NA,nrow(pars.wd25)),
                       beta=NA,
                       sd=NA)
  set.seed(10)
  for(i in 1:nrow(pars.wd25)){
    tempdat=rlnorm(n=length(tdat$meanHarvR[tdat$treat=="wd25"]),
                   meanlog = pars.wd25[i,1],
                   sdlog = pars.wd25[i,2])
    pval.wd25$alpha[i]=pars.wd25[i,1]
    pval.wd25$beta[i]=pars.wd25[i,2]
    pval.wd25$sd[i]=sd(log(tempdat))
  }

  # now calculate the number of times the sd exceeds that of the real data

  pval.wd25$sdExceed=0

  wd25.sd=sd(log(tdat$meanHarvR[tdat$treat=="wd25"]))

  pval.wd25$sdExceed[pval.wd25$sd>wd25.sd]=1

  #### Pb WD50 ####
  pval.wd50=data.frame(alpha=rep(NA,nrow(pars.wd50)),
                       beta=NA,
                       sd=NA)
  set.seed(10)
  for(i in 1:nrow(pars.wd50)){
    tempdat=rlnorm(n=length(tdat$meanHarvR[tdat$treat=="wd50"]),
                   meanlog = pars.wd50[i,1],
                   sdlog = pars.wd50[i,2])
    pval.wd50$alpha[i]=pars.wd50[i,1]
    pval.wd50$beta[i]=pars.wd50[i,2]
    pval.wd50$sd[i]=sd(log(tempdat))
  }

  # now calculate the number of times the sd exceeds that of the real data

  pval.wd50$sdExceed=0

  wd50.sd=sd(log(tdat$meanHarvR[tdat$treat=="wd50"]))

  pval.wd50$sdExceed[pval.wd50$sd>wd50.sd]=1

  #### Pb WE50 ####
  pval.we50=data.frame(alpha=rep(NA,nrow(pars.we50)),
                       beta=NA,
                       sd=NA)
  set.seed(10)
  for(i in 1:nrow(pars.we50)){
    tempdat=rlnorm(n=length(tdat$meanHarvR[tdat$treat=="we50"]),
                   meanlog = pars.we50[i,1],
                   sdlog = pars.we50[i,2])
    pval.we50$alpha[i]=pars.we50[i,1]
    pval.we50$beta[i]=pars.we50[i,2]
    pval.we50$sd[i]=sd(log(tempdat))
  }

  # now calculate the number of times the sd exceeds that of the real data

  pval.we50$sdExceed=0

  we50.sd=sd(log(tdat$meanHarvR[tdat$treat=="we50"]))

  pval.we50$sdExceed[pval.we50$sd>we50.sd]=1

  #df to hold pvals for model comparison to self, a way of knowing the model fit the data well
  t.pself=data.frame(year=rep(loopY[y],6),
                     scenario=c("Actual", "No Winter", "May-August","25% weeday removal","50% weekday removal","50% weekend removal"),
                     sd.pval=NA)
  # adding self comparison pvals
  
  t.pself$sd.pval=c(sum(pval.actual$sdExceed)/nrow(pval.actual),
                          sum(pval.noWinter$sdExceed)/nrow(pval.noWinter),
                          sum(pval.mayAug$sdExceed)/nrow(pval.mayAug),
                          sum(pval.wd25$sdExceed)/nrow(pval.wd25),
                          sum(pval.wd50$sdExceed)/nrow(pval.wd50),
                          sum(pval.we50$sdExceed)/nrow(pval.we50))
  bpval.self.y=rbind(bpval.self.y,t.pself)
  ## p-value tests comparing the scenarios to actual to show that some scenarios approximate the actual quite well.

  t.pcomp=data.frame(year=rep(loopY[y],6),
                     scenario=c("Actual", "No Winter", "May-August","25% weeday removal","50% weekday removal","50% weekend removal"),
                     sd.pval=NA)

  pval.actual$sdComp=0
  pval.actual$sdComp[pval.actual$sd>actual.sd]=1

  pval.noWinter$sdComp=0
  pval.noWinter$sdComp[pval.noWinter$sd>actual.sd]=1

  pval.mayAug$sdComp=0
  pval.mayAug$sdComp[pval.mayAug$sd>actual.sd]=1

  pval.wd25$sdComp=0
  pval.wd25$sdComp[pval.wd25$sd>actual.sd]=1

  pval.wd50$sdComp=0
  pval.wd50$sdComp[pval.wd50$sd>actual.sd]=1

  pval.we50$sdComp=0
  pval.we50$sdComp[pval.we50$sd>actual.sd]=1

  t.pcomp$sd.pval=c(sum(pval.actual$sdComp)/nrow(pval.actual),
                       sum(pval.noWinter$sdComp)/nrow(pval.noWinter),
                       sum(pval.mayAug$sdComp)/nrow(pval.mayAug),
                       sum(pval.wd25$sdComp)/nrow(pval.wd25),
                       sum(pval.wd50$sdComp)/nrow(pval.wd50),
                       sum(pval.we50$sdComp)/nrow(pval.we50))
  bpval.comp.y=rbind(bpval.comp.y,t.pcomp)
  
  # adding GR results to the output dataframe
  t.gr=data.frame(year=rep(loopY[y],6),
                  scenario=c("Actual", "No Winter", "May-August","25% weeday removal","50% weekday removal","50% weekend removal"),
                  gr.prsf=c(gr.a[[2]],gr.nw[[2]],gr.ma[[2]],gr.wd25[[2]],gr.wd50[[2]],gr.we50[[2]]),
                  gr.par1=c(gr.a[[1]][1,1],gr.nw[[1]][1,1],gr.ma[[1]][1,1],gr.wd25[[1]][1,1],gr.wd50[[1]][1,1],gr.we50[[1]][1,1]),
                  gr.par2=c(gr.a[[1]][2,1],gr.nw[[1]][2,1],gr.ma[[1]][2,1],gr.wd25[[1]][2,1],gr.wd50[[1]][2,1],gr.we50[[1]][2,1]))
  grMetrics.y=rbind(grMetrics.y,t.gr)
}

# saving big loop's output since it takes a while to run
yearLoopOutput=list(bpval.self.y,bpval.comp.y,grMetrics.y)
#saveRDS(yearLoopOutput, file = "yearLoopOutput_harvestR_9.29.25.RData")


```

```{r year-loop-preload, fig.width=8, fig.height=8, fig.cap="Distribution of harvest across years in the creel data set. Different data reduction scenarios are noted with varying colors."}
# first a couple exploratory plots of u distribution by year
# reading in model object from above because the loop takes a while to run
yearLoopOutput=readRDS("yearLoopOutput_harvestR_9.29.25.RData")

bpval.self.y=yearLoopOutput[[1]]
bpval.self.y=bpval.self.y[!is.na(bpval.self.y$scenario),]
bpval.comp.y=yearLoopOutput[[2]]
bpval.comp.y=bpval.comp.y[!is.na(bpval.comp.y$scenario),]
grMetrics.y=yearLoopOutput[[3]]
grMetrics.y=grMetrics.y[!is.na(grMetrics.y$scenario),]
ggplot(ttrExp)+theme_classic()+
  geom_density(aes(log(meanHarvR),fill=treat),alpha=0.2)+
  facet_wrap(~year,scales = 'free_y')+scale_fill_viridis_d()+
  theme(legend.position = "bottom")+
  labs(y="Density",x="Log(Harvest Rate)", fill="Scenario")
```

```{r year-loop-model-fit, fig.cap="Comparison of coefficient of variations for bayesian p-values comparing the actual data to the simulated data for the creel year and data reduction scenario."}

ggplot(bpval.self.y)+theme_classic()+
  geom_point(aes(x=year, y=sd.pval))+facet_wrap(~scenario)+
  geom_hline(yintercept = c(0.9,0.5,0.1),color="red")+
  coord_cartesian(ylim=c(0,1))
```

```{r yearLoopComparisonToActual, fig.cap="Comparison of the coeffienct of variations for bayesian p-values comparing the simulated data from each creel year and data reduction scenario to the actual data for that same creel year."}

ggplot(bpval.comp.y)+theme_classic()+
  geom_point(aes(x=year, y=sd.pval))+facet_wrap(~scenario)+
  geom_hline(yintercept = c(0.9,0.5,0.1),color="red")+
  coord_cartesian(ylim=c(0,1))
```
```{r yearByyearConvergence, fig.cap="Gelman-Rubin diagnostic results for the year-by-year model fitting. All years with  available data  (more than one non-zero harvest) had models successfully converge with a Gelman-Rubin score below the 1.1 threshold required for convergence (red horizonal line)."}
# model convergence
ggplot(grMetrics.y)+theme_classic()+
  geom_point(aes(x=year, y=gr.prsf))+facet_wrap(~scenario)+
  coord_cartesian(ylim=c(1,1.2))+
  geom_hline(yintercept = 1.1, color='red')+
  labs(y='Gelman-Rubin Score',x='Year')
```

### Alternative modeling distributions

Instead of the lognormal distribution there are two other probability distributions that could be used to model this data based on the characteristics of the data and the data characteristics assumed by each of the distributions.

Table of data distributions and their definitions:

| Distribution | Definition                                                                                                                                |
|-------------------------------------|-----------------------------------|
| Lognormal    | Continuously distributed quantities with nonnegative values. Random variables with the property that their logs are normally distributed. |
| Gamma        | Any continuous quantity that is nonnegative. Continuous version of a Poisson distribution.                                                |
| Beta         | Continuous random variable than can take on values between 0 and 1. Any random variable that can be expressed as a proportion.            |

The following analyses will compare the ability of models using the lognormal and beta distributions to fit the full data and each of the reduced data set.
A cursory look at the ability of gamma distributed models showed no real difference from lognormal and the definition of the beta better represents the data we're dealing with than the gamma definition so further analyses with this family of models was not pursued (Figure \@ref(fig:quickBetaFit)).
The purpose of the comparison between lognormal and beta family models is not to see which model provided the most convenient answer but to compare their ability to fit the data.
All comparisons presented in this analysis are comparing a model's ability to fit the data **within** each dataset and **not** across data sets which is what is necessary to gain inference on the effect of any data reductions.

```{r betaFits}
## FITTING ALTERNATIVE DISTRIBUTIONS ####
# combining actual and reduced dataframes into one big one for plotting

ttrExp=rbind(cbind(harv.surv.a,treat=rep("actual",nrow(harv.surv.a))),
             cbind(harv.surv.nw, treat=rep("noWinter",nrow(harv.surv.nw))),
             cbind(harv.surv.ma, treat=rep("mayAug",nrow(harv.surv.ma))),
             cbind(harv.surv.25wd, treat=rep("wd25",nrow(harv.surv.25wd))),
             cbind(harv.surv.50wd, treat=rep("wd50",nrow(harv.surv.50wd))),
             cbind(harv.surv.50we, treat=rep("we50",nrow(harv.surv.50we))))

# removing 0s, not necessary for gamma fit, but is for lognormal fit
ttrExp=ttrExp[ttrExp$meanHarvR!=0 & !is.na(ttrExp$meanHarvR),]
ggplot(ttrExp)+theme_classic()+
  geom_density(aes(x=meanHarvR, fill=treat),alpha=0.2)

harvLL.a.gamma=function(param){
  shape=param[1]
  rate=param[2]

  hrv=rgamma(nrow(ttrExp[ttrExp$treat=="actual",]), shape = shape, rate = rate)
  mu=mean(hrv,na.rm = T)
  sigma2=var(hrv,na.rm = T)
  
  ll=dgamma(ttrExp$meanHarvR[ttrExp$treat=="actual"], shape = mu^2/sigma2, rate = mu/sigma2, log = T)
  return(sum(ll))
}

prior.gamma=createUniformPrior(lower=c(0,0), upper = c(5,50))
setup.actual=createBayesianSetup(harvLL.a.gamma, prior = prior.gamma)

settings=list(iterations=20000, nrChains=3, message=F, burnin=5000)

startT=Sys.time()
set.seed(10)
harvR.actual.gamma=runMCMC(bayesianSetup = setup.actual, sampler = "DEzs", settings = settings)
endT=Sys.time() # takes about 10 sec

harvLL.a.beta=function(param){
  alpha=param[1]
  beta=param[2]

  us=rbeta(nrow(ttrExp[ttrExp$treat=="actual",]), shape1 = alpha, shape2 = beta)
  mu=mean(us)
  sigma2=var(us)
  
  ll=dbeta(ttrExp$meanHarvR[ttrExp$treat=="actual"], shape1 = ((mu^2)-(mu^3)-(mu*sigma2))/sigma2, shape2 = (mu-(2*(mu^2))+(mu^3)-sigma2+(mu*sigma2))/sigma2, log = T)
  return(sum(ll))
}

prior.beta=createUniformPrior(lower=c(0,0), upper = c(50,50))
setup.actual=createBayesianSetup(harvLL.a.beta, prior = prior.beta)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)

startT=Sys.time()
set.seed(10)
harvR.actual.beta=runMCMC(bayesianSetup = setup.actual, sampler = "DEzs", settings = settings)
endT=Sys.time() # takes about 1 sec

harvLL.a=function(param){
  alpha=param[1]
  beta=param[2]

  hrv=rlnorm(nrow(ttrExp[ttrExp$treat=="actual",]),meanlog = alpha, sdlog = beta)
  ll=dlnorm(ttrExp$meanHarvR[ttrExp$treat=="actual"], meanlog = mean(log(hrv)), sdlog = sd(log(hrv)), log = T)
  return(sum(ll))
}

prior=createTruncatedNormalPrior(mean=c(mean(log(ttrExp$meanHarvR[ttrExp$treat=="actual"])),sd(log(ttrExp$meanHarvR[ttrExp$treat=="actual"]))),
                                 sd=c(1,1),
                                 lower = c(-15,0),
                                 upper = c(15,5))

setup.actual=createBayesianSetup(harvLL.a, prior = prior)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
set.seed(10)
harvR.actual=runMCMC(bayesianSetup = setup.actual, sampler = "DEzs", settings = settings) # takes about 10 seconds


# working through rest of the reduced data sets and then calculating pvalues
#### NW ####
harvLL.nw.beta=function(param){
  alpha=param[1]
  beta=param[2]

  us=rbeta(nrow(ttrExp[ttrExp$treat=="noWinter",]), shape1 = alpha, shape2 = beta)
  mu=mean(us)
  sigma2=var(us)
  
  ll=dbeta(ttrExp$meanHarvR[ttrExp$treat=="noWinter"], shape1 = ((mu^2)-(mu^3)-(mu*sigma2))/sigma2, shape2 = (mu-(2*(mu^2))+(mu^3)-sigma2+(mu*sigma2))/sigma2, log = T)
  return(sum(ll))
}

prior.beta=createUniformPrior(lower=c(0,0), upper = c(50,50))
setup.nw.beta=createBayesianSetup(harvLL.nw.beta, prior = prior.beta)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)

set.seed(10)
harvR.nw.beta=runMCMC(bayesianSetup = setup.nw.beta, sampler = "DEzs", settings = settings)

#### MA ####

harvLL.ma.beta=function(param){
  alpha=param[1]
  beta=param[2]

  us=rbeta(nrow(ttrExp[ttrExp$treat=="mayAug",]), shape1 = alpha, shape2 = beta)
  mu=mean(us)
  sigma2=var(us)
  
  ll=dbeta(ttrExp$meanHarvR[ttrExp$treat=="mayAug"], shape1 = ((mu^2)-(mu^3)-(mu*sigma2))/sigma2, shape2 = (mu-(2*(mu^2))+(mu^3)-sigma2+(mu*sigma2))/sigma2, log = T)
  return(sum(ll))
}

prior.beta=createUniformPrior(lower=c(0,0), upper = c(50,50))
setup.ma.beta=createBayesianSetup(harvLL.ma.beta, prior = prior.beta)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)

set.seed(10)
harvR.ma.beta=runMCMC(bayesianSetup = setup.ma.beta, sampler = "DEzs", settings = settings)

#### WD25 ####

harvLL.wd25.beta=function(param){
  alpha=param[1]
  beta=param[2]

  us=rbeta(nrow(ttrExp[ttrExp$treat=="wd25",]), shape1 = alpha, shape2 = beta)
  mu=mean(us)
  sigma2=var(us)
  
  ll=dbeta(ttrExp$meanHarvR[ttrExp$treat=="wd25"], shape1 = ((mu^2)-(mu^3)-(mu*sigma2))/sigma2, shape2 = (mu-(2*(mu^2))+(mu^3)-sigma2+(mu*sigma2))/sigma2, log = T)
  return(sum(ll))
}

prior.beta=createUniformPrior(lower=c(0,0), upper = c(50,50))
setup.wd25.beta=createBayesianSetup(harvLL.wd25.beta, prior = prior.beta)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)
set.seed(10)
harvR.25wd.beta=runMCMC(bayesianSetup = setup.wd25.beta, sampler = "DEzs", settings = settings)

#### WD50 ####
harvLL.wd50.beta=function(param){
  alpha=param[1]
  beta=param[2]

  us=rbeta(nrow(ttrExp[ttrExp$treat=="wd50",]), shape1 = alpha, shape2 = beta)
  mu=mean(us)
  sigma2=var(us)
  
  ll=dbeta(ttrExp$meanHarvR[ttrExp$treat=="wd50"], shape1 = ((mu^2)-(mu^3)-(mu*sigma2))/sigma2, shape2 = (mu-(2*(mu^2))+(mu^3)-sigma2+(mu*sigma2))/sigma2, log = T)
  return(sum(ll))
}

prior.beta=createUniformPrior(lower=c(0,0), upper = c(50,50))
setup.wd50.beta=createBayesianSetup(harvLL.wd50.beta, prior = prior.beta)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)

set.seed(10)
harvR.50wd.beta=runMCMC(bayesianSetup = setup.wd50.beta, sampler = "DEzs", settings = settings)

#### WE50 ####
harvLL.we50.beta=function(param){
  alpha=param[1]
  beta=param[2]

  us=rbeta(nrow(ttrExp[ttrExp$treat=="actual",]), shape1 = alpha, shape2 = beta)
  mu=mean(us)
  sigma2=var(us)
  
  ll=dbeta(ttrExp$meanHarvR[ttrExp$treat=="actual"], shape1 = ((mu^2)-(mu^3)-(mu*sigma2))/sigma2, shape2 = (mu-(2*(mu^2))+(mu^3)-sigma2+(mu*sigma2))/sigma2, log = T)
  return(sum(ll))
}

prior.beta=createUniformPrior(lower=c(0,0), upper = c(50,50))
setup.we50.beta=createBayesianSetup(harvLL.we50.beta, prior = prior.beta)

settings=list(iterations=10000, nrChains=3, message=F, burnin=5000)

set.seed(10)
harvR.50we.beta=runMCMC(bayesianSetup = setup.we50.beta, sampler = "DEzs", settings = settings)
```

```{r quickBetaFit, fig.cap="Comparison of model fits for lognormal, gamma, and beta family of models. This rough look at the models suggests there are no obvious differences between the choice of modeling distribution. This will be evaluated statistically using bayesian p values later on."}
# evaluating model fit

pars.a.lnorm=getSample(harvR.actual)
pars.a.gamma=getSample(harvR.actual.gamma)
pars.a.beta=getSample(harvR.actual.beta)

set.seed(10)
aComp.lnorm=data.frame(harvR=c(ttrExp$meanHarvR[ttrExp$treat=="actual"],rlnorm(n=length(ttrExp$meanHarvR[ttrExp$treat=="actual"]),
                                                                    meanlog = median(pars.a.lnorm[,1]),
                                                                    sdlog = median(pars.a.lnorm[,2]))),
                 treat=c(rep("observed",length(ttrExp$meanHarvR[ttrExp$treat=="actual"])),rep("pred",length(ttrExp$meanHarvR[ttrExp$treat=="actual"]))))

aComp.gamma=data.frame(harvR=c(ttrExp$meanHarvR[ttrExp$treat=="actual"],rgamma(n=length(ttrExp$meanHarvR[ttrExp$treat=="actual"]),
                                                                    shape = median(pars.a.gamma[,1]),
                                                                    rate = median(pars.a.gamma[,2]))),
                  treat=c(rep("observed",length(ttrExp$meanHarvR[ttrExp$treat=="actual"])),rep("pred",length(ttrExp$meanHarvR[ttrExp$treat=="actual"]))))

aComp.beta=data.frame(harvR=c(ttrExp$meanHarvR[ttrExp$treat=="actual"],rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="actual"]),
                                                                    shape1 = median(pars.a.beta[,1]),
                                                                    shape2 = median(pars.a.beta[,2]))),
                  treat=c(rep("observed",length(ttrExp$meanHarvR[ttrExp$treat=="actual"])),rep("pred",length(ttrExp$meanHarvR[ttrExp$treat=="actual"]))))

ln=ggplot(aComp.lnorm)+theme_classic()+
  geom_density(aes(x=harvR, fill=treat),alpha=0.3)+
  scale_fill_viridis_d()+labs(x="Harvest Rate",title = "Lognormal")+
  theme(legend.position = c(0.75,0.75))
gma=ggplot(aComp.gamma)+theme_classic()+
  geom_density(aes(x=harvR, fill=treat),alpha=0.3)+
  scale_fill_viridis_d()+labs(x="Harvest Rate",title = "Gamma")+
  theme(legend.position = c(0.75,0.75))
bta=ggplot(aComp.beta)+theme_classic()+
  geom_density(aes(x=harvR, fill=treat),alpha=0.3)+
  scale_fill_viridis_d()+labs(x="Harvest Rate",title = "Beta")+
  theme(legend.position = c(0.75,0.75))
ggarrange(ln,gma,bta)

```

```{r betaFitViz, fig.cap="Visualization of the model fit to the data for each scenario using a beta distribution instead of a lognormal distribution."}
# looking to see if parm estimates produce data that visually at least looks like the observed data for that scenario
pars.a.beta=getSample(harvR.actual.beta)
pars.nw.beta=getSample(harvR.nw.beta)
pars.ma.beta=getSample(harvR.ma.beta)
pars.wd25.beta=getSample(harvR.25wd.beta)
pars.wd50.beta=getSample(harvR.50wd.beta)
pars.we50.beta=getSample(harvR.50we.beta)
set.seed(10)
aComp.beta=data.frame(harvR=c(ttrExp$meanHarvR[ttrExp$treat=="actual"],rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="actual"]),
                                                                    shape1 = median(pars.a.beta[,1]),
                                                                    shape2 = median(pars.a.beta[,2]))),
                  treat=c(rep("observed",length(ttrExp$meanHarvR[ttrExp$treat=="actual"])),rep("pred",length(ttrExp$meanHarvR[ttrExp$treat=="actual"]))))
nwComp.beta=data.frame(harvR=c(ttrExp$meanHarvR[ttrExp$treat=="noWinter"],rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="noWinter"]),
                                                                    shape1 = median(pars.nw.beta[,1]),
                                                                    shape2 = median(pars.nw.beta[,2]))),
                  treat=c(rep("observed",length(ttrExp$meanHarvR[ttrExp$treat=="noWinter"])),rep("pred",length(ttrExp$meanHarvR[ttrExp$treat=="noWinter"]))))

maComp.beta=data.frame(harvR=c(ttrExp$meanHarvR[ttrExp$treat=="mayAug"],rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="mayAug"]),
                                                                    shape1 = median(pars.ma.beta[,1]),
                                                                    shape2 = median(pars.ma.beta[,2]))),
                  treat=c(rep("observed",length(ttrExp$meanHarvR[ttrExp$treat=="mayAug"])),rep("pred",length(ttrExp$meanHarvR[ttrExp$treat=="mayAug"]))))

wd25Comp.beta=data.frame(harvR=c(ttrExp$meanHarvR[ttrExp$treat=="wd25"],rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="wd25"]),
                                                                    shape1 = median(pars.wd25.beta[,1]),
                                                                    shape2 = median(pars.wd25.beta[,2]))),
                    treat=c(rep("observed",length(ttrExp$meanHarvR[ttrExp$treat=="wd25"])),rep("pred",length(ttrExp$meanHarvR[ttrExp$treat=="wd25"]))))

wd50Comp.beta=data.frame(harvR=c(ttrExp$meanHarvR[ttrExp$treat=="wd50"],rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="wd50"]),
                                                                    shape1 = median(pars.wd50.beta[,1]),
                                                                    shape2 = median(pars.wd50.beta[,2]))),
                    treat=c(rep("observed",length(ttrExp$meanHarvR[ttrExp$treat=="wd50"])),rep("pred",length(ttrExp$meanHarvR[ttrExp$treat=="wd50"]))))

we50Comp.beta=data.frame(harvR=c(ttrExp$meanHarvR[ttrExp$treat=="we50"],rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="we50"]),
                                                                    shape1 = median(pars.we50.beta[,1]),
                                                                    shape2 = median(pars.we50.beta[,2]))),
                    treat=c(rep("observed",length(ttrExp$meanHarvR[ttrExp$treat=="we50"])),rep("pred",length(ttrExp$meanHarvR[ttrExp$treat=="we50"]))))

a.p.beta=ggplot(aComp.beta)+theme_classic()+
  geom_density(aes(x=harvR,fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Harvest Rate", y="Density", title = "Actual", fill=element_blank())
nw.p.beta=ggplot(nwComp.beta)+theme_classic()+
  geom_density(aes(x=harvR,fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Harvest Rate", y="Density", title = "No Winter Data \n(April - October)", fill=element_blank())
ma.p.beta=ggplot(maComp.beta)+theme_classic()+
  geom_density(aes(x=harvR,fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Harvest Rate", y="Density", title = "May - August \nData Only", fill=element_blank())
wd25.p.beta=ggplot(wd25Comp.beta)+theme_classic()+
  geom_density(aes(x=harvR,fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Harvest Rate", y="Density", title = "25% of Weekday \nCreel Visits/Month \nRemoved", fill=element_blank())
wd50.p.beta=ggplot(wd25Comp.beta)+theme_classic()+
  geom_density(aes(x=harvR,fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Harvest Rate", y="Density", title = "50% of Weekday \nCreel Visits/Month \nRemoved", fill=element_blank())
we50.p.beta=ggplot(we50Comp.beta)+theme_classic()+
  geom_density(aes(x=harvR,fill=treat),alpha=0.2)+
  scale_fill_viridis_d()+labs(x="Harvest Rate", y="Density", title = "50% of Weekdend \nCreel Visits/Month \nRemoved", fill=element_blank())

ggarrange(a.p.beta,nw.p.beta,ma.p.beta,wd25.p.beta,wd50.p.beta,we50.p.beta, common.legend = T)
```

```{r betaGRDiagnostics, eval=F, echo=T}
### MODEL CHECKNG ####

gelmanDiagnostics(harvR.actual.beta) # converged
gelmanDiagnostics(harvR.nw.beta) # converged
gelmanDiagnostics(harvR.ma.beta) # converged
gelmanDiagnostics(harvR.25wd.beta) # converged
gelmanDiagnostics(harvR.50wd.beta) # converged
gelmanDiagnostics(harvR.50we.beta) # converged

```

```{r betaPVals}
# bayesian p-value for each model
# dataframe to hold output
bpval.self.beta=data.frame(scenario=c("Actual", "No Winter", "May-August","25% weeday removal","50% weekday removal","50% weekend removal"),
                    sd.pval=NA)
#### Pb ACTUAL ####
pval.actual=data.frame(shape1=rep(NA,nrow(pars.a.beta)),
                       shape2=NA,
                       sd=NA)
set.seed(10)
for(i in 1:nrow(pars.a.beta)){
  tempdat=rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="actual"]),
                 shape1 = pars.a.beta[i,1],
                 shape2 = pars.a.beta[i,2])
  pval.actual$shape1[i]=pars.a.beta[i,1]
  pval.actual$shape2[i]=pars.a.beta[i,2]
  pval.actual$sd[i]=sd(tempdat)
}

# now calculate the number of times the sd exceeds that of the real data

pval.actual$sdExceed=0

actual.sd=sd(ttrExp$meanHarvR[ttrExp$treat=="actual"])

pval.actual$sdExceed[pval.actual$sd>actual.sd]=1

bpval.self.beta$sd.pval[1]=sum(pval.actual$sdExceed==1)/nrow(pval.actual) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal


#### Pb NO WINTER ####
pval.noWinter=data.frame(shape1=rep(NA,nrow(pars.nw.beta)),
                         shape2=NA,
                         sd=NA)
set.seed(10)
for(i in 1:nrow(pars.nw.beta)){
  tempdat=rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="noWinter"]),
                shape1 = pars.nw.beta[i,1],
                shape2 = pars.nw.beta[i,2])
  pval.noWinter$shape1[i]=pars.nw.beta[i,1]
  pval.noWinter$shape2[i]=pars.nw.beta[i,2]
  pval.noWinter$sd[i]=sd(tempdat)
}

# now calculate the number of times the sd exceeds that of the real data

pval.noWinter$sdExceed=0

noWinter.sd=sd(ttrExp$meanHarvR[ttrExp$treat=="noWinter"])

pval.noWinter$sdExceed[pval.noWinter$sd>noWinter.sd]=1

bpval.self.beta$sd.pval[2]=sum(pval.noWinter$sdExceed==1)/nrow(pval.noWinter) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

#### Pb MAYAUGUST ####
pval.mayAug=data.frame(shape1=rep(NA,nrow(pars.ma.beta)),
                       shape2=NA,
                       sd=NA)
set.seed(10)
for(i in 1:nrow(pars.ma.beta)){
  tempdat=rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="mayAug"]),
                shape1 = pars.ma.beta[i,1],
                shape2 = pars.ma.beta[i,2])
  pval.mayAug$shape1[i]=pars.ma.beta[i,1]
  pval.mayAug$shape2[i]=pars.ma.beta[i,2]
  pval.mayAug$sd[i]=sd(tempdat)
}

# now calculate the number of times the sd exceeds that of the real data

pval.mayAug$sdExceed=0

mayAug.sd=sd(ttrExp$meanHarvR[ttrExp$treat=="mayAug"])

pval.mayAug$sdExceed[pval.mayAug$sd>mayAug.sd]=1

bpval.self.beta$sd.pval[3]=sum(pval.mayAug$sdExceed==1)/nrow(pval.mayAug) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

#### Pb WD25 ####
pval.wd25=data.frame(shape1=rep(NA,nrow(pars.wd25.beta)),
                     shape2=NA,
                     sd=NA)
set.seed(10)
for(i in 1:nrow(pars.wd25.beta)){
  tempdat=rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="wd25"]),
                shape1 = pars.wd25.beta[i,1],
                shape2 = pars.wd25.beta[i,2])
  pval.wd25$shape1[i]=pars.wd25.beta[i,1]
  pval.wd25$shape2[i]=pars.wd25.beta[i,2]
  pval.wd25$sd[i]=sd(tempdat)
}

# now calculate the number of times the sd exceeds that of the real data

pval.wd25$sdExceed=0

wd25.sd=sd(ttrExp$meanHarvR[ttrExp$treat=="wd25"])

pval.wd25$sdExceed[pval.wd25$sd>wd25.sd]=1

bpval.self.beta$sd.pval[4]=sum(pval.wd25$sdExceed==1)/nrow(pval.wd25) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

#### Pb WD50 ####
pval.wd50=data.frame(shape1=rep(NA,nrow(pars.wd50.beta)),
                     shape2=NA,
                     sd=NA)
set.seed(10)
for(i in 1:nrow(pars.wd50.beta)){
  tempdat=rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="wd50"]),
                shape1 = pars.wd50.beta[i,1],
                shape2 = pars.wd50.beta[i,2])
  pval.wd50$shape1[i]=pars.wd50.beta[i,1]
  pval.wd50$shape2[i]=pars.wd50.beta[i,2]
  pval.wd50$sd[i]=sd(tempdat)
}

# now calculate the number of times the sd exceeds that of the real data

pval.wd50$sdExceed=0

wd50.sd=sd(ttrExp$meanHarvR[ttrExp$treat=="wd50"])

pval.wd50$sdExceed[pval.wd50$sd>wd50.sd]=1

bpval.self.beta$sd.pval[5]=sum(pval.wd50$sdExceed==1)/nrow(pval.wd50) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

#### Pb WE50 ####
pval.we50=data.frame(shape1=rep(NA,nrow(pars.we50.beta)),
                     shape2=NA,
                     sd=NA)
set.seed(10)
for(i in 1:nrow(pars.we50.beta)){
  tempdat=rbeta(n=length(ttrExp$meanHarvR[ttrExp$treat=="we50"]),
                shape1 = pars.we50.beta[i,1],
                shape2 = pars.we50.beta[i,2])
  pval.we50$shape1[i]=pars.we50.beta[i,1]
  pval.we50$shape2[i]=pars.we50.beta[i,2]
  pval.we50$sd[i]=sd(tempdat)
}

# now calculate the number of times the sd exceeds that of the real data

pval.we50$sdExceed=0

we50.sd=sd(ttrExp$meanHarvR[ttrExp$treat=="we50"])

pval.we50$sdExceed[pval.we50$sd>we50.sd]=1

bpval.self.beta$sd.pval[6]=sum(pval.we50$sdExceed==1)/nrow(pval.we50) # if this value is < 0.1 for >0.9 that would suggest model is not fitting well. 0.5 is ideal

kable(bpval.self.beta, digits = 3, col.names = c("Scenario", "SD p-value"), align="lcc",caption = "Bayesian p-values for a beta distributed model. Each p-value describes whether or not there is a significant difference between data generated by the fitted model and the actual data for that scenario. Standard devidation (SD) was the metric assessed here.")


```

The Bayesian p-values in the table presented here suggests the beta models do a poor job of reproducing the data they were fit to (Table \@ref(tab:betaPVals)).
This inference comes from the lack of p-values between 0.1 - 0.9 for the CV and SD metrics.
Because this model is unable to reproduce the data it was fit to, we know that this model is not a useful one for fitting these data.
This means that the beta distribution is not likely to be appropriate for making comparisons between the actual data and the reduced data to understand whether or not creel effort reductions result in significantly different harvest rates or not.

### P-value sensitivity analysis


#### Interpreting each statistic

Some more detail on what I'm looking for in each statistic and what the values mean that are used to calculate each Bayesian p-value.
It's important to remember that the way the sampling algorithm works in this Bayesian framework the frequency of values in the MCMC output is relative to their support in the data.
In other words, parameter values that produce better models fits show up in the chain more often.
This is critical to understanding the Bayesian p-values here because each test statistic is calculated for simulated data based on the MCMC output so each test statistic value should appear with a frequency equal to the support provided by the data too.
For example, the medians calculated from data sets produced by each parameter set in the MCMC output should produce medians that have frequencies equal to how well those parameters fit the data.
Thus medians well below the median of the observed data should be pretty rare (and likewise for medians well above) because the parameter set from the MCMC chain that produces data with said median should be relatively infrequent in the MCMC output.
Medians close to the median of the observed data should be more common, if the model is fitting the data well, because the parameters generating those means are more common in the MCMC output due to the better fit to the data they provide.
This results in medians that are often close to the median of the observed data and by random chance should be just as likely to be above as below the median of the observed data.
This is why Bayesian p-values close to 0.5 signify a good model fit, because about 50% of the time the test statistic is more extreme than the value for the observed data.
If this values is extremely high or low then is signals that our model is not adequately representing the observed data.

##### Standard Deviation

Another straightforward calculation.
SD is calculated for a data set drawn from a random lognormal distribution parameterized using the values in the MCMC chain.
This is repeated for each set of parameter values in the MCMC chain to create as many simulated data sets as there are iterations in the MCMC chain.
The SD of each of these simulated data sets is compared to the SD for the observed data for that scenario (Actual, No Winter, May - August, etc.).
Whether the SD of the simulated data is greater than or less than the SD of the observed data is recorded.
**If the model is able to reproduce the observed data well then the SD of the simulated data should be greater than the SD of the observed data about 50% of the time and less than the SD of the observed data about 50% of the time too.**

##### Median

Simple descriptive statistic here.
The median is calculated for each simulated data set produced by drawing values from a random lognormal distribution parameterized using the values in the MCMC output.
The number of times the median of the simulated data set exceeds the median of the observed data is calculated and that proportion of times the test statistic exceeds the value for the observed data is the Bayesian p-value.
**If the model is adequately representing the data then the medians generated from the simulated data sets should exceed the median of the observed data roughly 50% of the time.**

##### Kurtosis

This metric essentially measures the shape of a distribution, specifically the tails.
Kurtosis is calculated for the simulated data for each set of parameter values in the MCMC output and compared to the kurtosis of the observed data for the given scenario (Actual, No Winter, May - August, etc.).
When the kurtosis value of the simulated data is greater than the kurtosis of the observed data this signals more frequent extreme values in the data.
In other words the tails of the distribution of the simulated data are thicker than the tails of the distribution of the observed data.
The opposite is true when the kurtosis of the simulated data is less than the kurtosis of the observed data.
**A well-fitting model will produce data with kurtosis values similar to the the observed data's kurtosis. The kurtosis of the simulated data will exceed that of the observed data roughly 50% of the time. If the kurtosis of the simulated data is more frequently less than the observed data this would signal that the model is not doing well as representing the rare values in the tails of the data distribution. The opposite is true if kurtosis values for the simulated data tend to be larger than the observed data.**

##### Chi Square Test ($\chi^2$)

This test is performed outside the Bayesian p-value framework.
Instead, each simulated data set arising from a set of parameters in the MCMC output is compared to the probability distribution of the observed data to ask whether the simulated data arises from the same distribution as the observed data.
Again, because of the way the parameter space is sampled in MCMC algorithms the values that produce better model fits will be more common which means that in the $\chi^2$ test the null hypothesis that the data comes from the supplied probability distribution should be accepted more often than it's rejected.
**If the null hypothesis is rejected the majority of the time that would indicate that the model is not adequately representing the data if it can't produce simulated data that would appear to be from the probability distribution of the observed data used to fit the model in the first place.**

##### Fisher 'F' Statistic

The F statistic is the ratio of variances between the simulated and observed data sets.
For this test I'm looking to see if the F statistic for the simulated:observed comparison is greater than or less than 1 (the ratio of the variance for the observed:observed comparison).
**The closer the F statistic is to 1 for the simulated:observed comparison the better job the model is doing at adequately representing the data. So the number of times the F statistic exceeds 1 should, if the model is fitting the data well, be about 50% of the time. If the Bayesian p-value is much higher than 0.5 that would indicate that the variance of the simulated data is often greater than the variance of the observed data. The opposite then is true when the p-value is smaller than 0.5**

##### Coverage

The probability that the confidence interval for the parameter of interest will include the true value of the data. Here this means comparing confidence intervals on the estimated fishery metric calculated from the reduced data set back to the calculation of the fishery metric for the full creel data set. This is repeated for each parameter value in the MCMC chain resulting from the model fits. 
**The percentage of parameter values that produce fishery metric data that 'covers' what was observed in the full data set is what is reported. The higher the coverage, the better the reduced creel data set is able to approximate the full data set.**

```{r pStatTab}
tab=data.frame(statistic=c("SD","Median","Kurtosis","X2","F", "Coverage"),
               desription=c("Square root of the variance.",
                            "Middle of the data when all values are ordered from smallest to largest. Similar to a mean but less sensitive to outlier values.",
                            "Measure of the width, or 'tailedness' of a distribution. In other words, do the tails of the distribution contain more or fewer outliers (i.e. are they thicker or thinner)?",
                            "Chi square test statistic for a 'goodness of fit' test is calculated. Here I want to know if frequency of values in the model generated data come from the same distribution as the observed data",
                            "The ratio of the variance between the model simulated data and the observed data.",
                            "The probability that the 95 percent confidence interval for a paramter will contain the true value of the data."))
kable(tab,col.names=c('Statistic', 'Description'),align="cl",caption = "Candidate variance metrics to use for calculation of Bayesian p-values.")
```

```{r bpValCalcsAlternateMetrics}

# bayesian p-value for each model
pars.a=getSample(harv.actual)
pars.nw=getSample(harv.nw)
pars.ma=getSample(harv.ma)
pars.wd25=getSample(harv.25wd)
pars.wd50=getSample(harv.50wd)
pars.we50=getSample(harv.50we)

#### Pb ACTUAL ####
pval.actual=data.frame(alpha=rep(NA,nrow(pars.a)),
                       beta=NA,
                       sd=NA,
                       med=NA,
                       kurt=NA,
                       x2=NA,
                       f=NA,
                       lcl=NA,
                       ucl=NA)
set.seed(10)
for(i in 1:nrow(pars.a)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="actual"]),
                 meanlog = pars.a[i,1],
                 sdlog = pars.a[i,2])
  pval.actual$alpha[i]=pars.a[i,1]
  pval.actual$beta[i]=pars.a[i,2]
  pval.actual$sd[i]=sd(log(tempdat))
  pval.actual$med[i]=median(log(tempdat))
  pval.actual$kurt[i]=kurtosis(log(tempdat))
  ct=chisq.test(tempdat, p=modDat$meanHarvR[modDat$treat=="actual"], rescale.p = T)
  pval.actual$x2[i]=ct$p.value
  f.test=var.test(log(tempdat),log(modDat$meanHarvR[modDat$treat=="actual"]))
  pval.actual$f[i]=f.test$statistic
  cis=quantile(log(tempdat), probs=c(0.025, 0.975))
  pval.actual$lcl[i]=cis[1]
  pval.actual$ucl[i]=cis[2]
  
}

# now calculate the number of times the test statistic exceeds that of the real data

pval.actual$sdExceed=0
pval.actual$medExceed=0
pval.actual$kurtExceed=0
pval.actual$x2SigDiff=0
pval.actual$fStatExceed=0
pval.actual$coverageExceed=0

actual.sd=sd(log(modDat$meanHarvR[modDat$treat=="actual"]))
actual.med=median(log(modDat$meanHarvR[modDat$treat=="actual"]))
actual.kurt=kurtosis(log(modDat$meanHarvR[modDat$treat=="actual"]))
actual.x2=chisq.test(modDat$meanHarvR[modDat$treat=="actual"],p=modDat$meanHarvR[modDat$treat=="actual"],rescale.p = T)
actual.f=var.test(log(modDat$meanHarvR[modDat$treat=="actual"]),log(modDat$meanHarvR[modDat$treat=="actual"]))

pval.actual$sdExceed[pval.actual$sd>actual.sd]=1
pval.actual$medExceed[pval.actual$med>actual.med]=1
pval.actual$kurtExceed[pval.actual$kurt>actual.kurt]=1 # kurt value great that observed data means there are more outliers, or thicker tails than the observed data
pval.actual$x2SigDiff[pval.actual$x2<0.05]=1 # looking at how often we would say the simmed data did not come from the probability distribution of the actual data
pval.actual$fStatExceed[pval.actual$f>actual.f$statistic]=1 # how often the f statistic is bigger than the actual
pval.actual$coverageExceed[pval.actual$lcl<=mean(log(modDat$meanHarvR[modDat$treat=="actual"])) & pval.actual$ucl>=mean(log(modDat$meanHarvR[modDat$treat=="actual"]))]=1 # how often the cis from the temp data contain the mean of the actual data


#### Pb NoWinter ####
pval.noWinter=data.frame(alpha=rep(NA,nrow(pars.nw)),
                         beta=NA,
                         sd=NA,
                         med=NA,
                         kurt=NA,
                         x2=NA,
                         f=NA,
                         lcl=NA,
                         ucl=NA)
set.seed(10)
for(i in 1:nrow(pars.nw)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="noWinter"]),
                 meanlog = pars.nw[i,1],
                 sdlog = pars.nw[i,2])
  pval.noWinter$alpha[i]=pars.nw[i,1]
  pval.noWinter$beta[i]=pars.nw[i,2]
  pval.noWinter$sd[i]=sd(log(tempdat))
  pval.noWinter$med[i]=median(log(tempdat))
  pval.noWinter$kurt[i]=kurtosis(log(tempdat))
  ct=chisq.test(tempdat, p=modDat$meanHarvR[modDat$treat=="noWinter"], rescale.p = T)
  pval.noWinter$x2[i]=ct$p.value
  f.test=var.test(log(tempdat),log(modDat$meanHarvR[modDat$treat=="noWinter"]))
  pval.noWinter$f[i]=f.test$statistic
  cis=quantile(log(tempdat), probs=c(0.025, 0.975))
  pval.noWinter$lcl[i]=cis[1]
  pval.noWinter$ucl[i]=cis[2]
}

# now calculate the number of times the test stat exceeds that of the real data

pval.noWinter$sdExceed=0
pval.noWinter$medExceed=0
pval.noWinter$kurtExceed=0
pval.noWinter$x2SigDiff=0
pval.noWinter$fStatExceed=0
pval.noWinter$coverageExceed=0

noWinter.sd=sd(log(modDat$meanHarvR[modDat$treat=="noWinter"]))
noWinter.med=median(log(modDat$meanHarvR[modDat$treat=="noWinter"]))
noWinter.kurt=kurtosis(log(modDat$meanHarvR[modDat$treat=="noWinter"]))
noWinter.x2=chisq.test(modDat$meanHarvR[modDat$treat=="noWinter"],p=modDat$meanHarvR[modDat$treat=="noWinter"],rescale.p = T)
noWinter.f=var.test(log(modDat$meanHarvR[modDat$treat=="noWinter"]),log(modDat$meanHarvR[modDat$treat=="noWinter"]))

pval.noWinter$sdExceed[pval.noWinter$sd>noWinter.sd]=1
pval.noWinter$medExceed[pval.noWinter$med>noWinter.med]=1
pval.noWinter$kurtExceed[pval.noWinter$kurt>noWinter.kurt]=1 # kurt value great that observed data means there are more outliers, or thicker tails than the observed data
pval.noWinter$x2SigDiff[pval.noWinter$x2<0.05]=1 # looking at how often we would say the simmed data did not come from the probability distribution of the noWinter data
pval.noWinter$fStatExceed[pval.noWinter$f>noWinter.f$statistic]=1 # how often the f statistic is bigger than the noWinter
pval.noWinter$coverageExceed[pval.noWinter$lcl<=mean(log(modDat$meanHarvR[modDat$treat=="noWinter"])) & pval.noWinter$ucl>=mean(log(modDat$meanHarvR[modDat$treat=="noWinter"]))]=1 # how often the cis from the temp data contain the mean of the noWinter data

#### Pb MayAug ####
pval.mayAug=data.frame(alpha=rep(NA,nrow(pars.ma)),
                       beta=NA,
                       sd=NA,
                       med=NA,
                       kurt=NA,
                       x2=NA,
                       f=NA,
                       lcl=NA,
                       ucl=NA)
set.seed(10)
for(i in 1:nrow(pars.ma)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="mayAug"]),
                 meanlog = pars.ma[i,1],
                 sdlog = pars.ma[i,2])
  pval.mayAug$alpha[i]=pars.ma[i,1]
  pval.mayAug$beta[i]=pars.ma[i,2]
  pval.mayAug$sd[i]=sd(log(tempdat))
  pval.mayAug$med[i]=median(log(tempdat))
  pval.mayAug$kurt[i]=kurtosis(log(tempdat))
  ct=chisq.test(tempdat, p=modDat$meanHarvR[modDat$treat=="mayAug"], rescale.p = T)
  pval.mayAug$x2[i]=ct$p.value
  f.test=var.test(log(tempdat),log(modDat$meanHarvR[modDat$treat=="mayAug"]))
  pval.mayAug$f[i]=f.test$statistic
  cis=quantile(log(tempdat), probs=c(0.025, 0.975))
  pval.mayAug$lcl[i]=cis[1]
  pval.mayAug$ucl[i]=cis[2]
}

# now calculate the number of times the test stat exceeds that of the real data

pval.mayAug$sdExceed=0
pval.mayAug$medExceed=0
pval.mayAug$kurtExceed=0
pval.mayAug$x2SigDiff=0
pval.mayAug$fStatExceed=0
pval.mayAug$coverageExceed=0

mayAug.sd=sd(log(modDat$meanHarvR[modDat$treat=="mayAug"]))
mayAug.med=median(log(modDat$meanHarvR[modDat$treat=="mayAug"]))
mayAug.kurt=kurtosis(log(modDat$meanHarvR[modDat$treat=="mayAug"]))
mayAug.x2=chisq.test(modDat$meanHarvR[modDat$treat=="mayAug"],p=modDat$meanHarvR[modDat$treat=="mayAug"],rescale.p = T)
mayAug.f=var.test(log(modDat$meanHarvR[modDat$treat=="mayAug"]),log(modDat$meanHarvR[modDat$treat=="mayAug"]))

pval.mayAug$sdExceed[pval.mayAug$sd>mayAug.sd]=1
pval.mayAug$medExceed[pval.mayAug$med>mayAug.med]=1
pval.mayAug$kurtExceed[pval.mayAug$kurt>mayAug.kurt]=1 # kurt value great that observed data means there are more outliers, or thicker tails than the observed data
pval.mayAug$x2SigDiff[pval.mayAug$x2<0.05]=1 # looking at how often we would say the simmed data did not come from the probability distribution of the mayAug data
pval.mayAug$fStatExceed[pval.mayAug$f>mayAug.f$statistic]=1 # how often the f statistic is bigger than the mayAug
pval.mayAug$coverageExceed[pval.mayAug$lcl<=mean(log(modDat$meanHarvR[modDat$treat=="mayAug"])) & pval.mayAug$ucl>=mean(log(modDat$meanHarvR[modDat$treat=="mayAug"]))]=1 # how often the cis from the temp data contain the mean of the mayAug data

#### Pb WD25 ####
pval.wd25=data.frame(alpha=rep(NA,nrow(pars.wd25)),
                     beta=NA,
                     sd=NA,
                     med=NA,
                     kurt=NA,
                     x2=NA,
                     f=NA,
                     lcl=NA,
                     ucl=NA)
set.seed(10)
for(i in 1:nrow(pars.wd25)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="wd25"]),
                 meanlog = pars.wd25[i,1],
                 sdlog = pars.wd25[i,2])
  pval.wd25$alpha[i]=pars.wd25[i,1]
  pval.wd25$beta[i]=pars.wd25[i,2]
  pval.wd25$sd[i]=sd(log(tempdat))
  pval.wd25$med[i]=median(log(tempdat))
  pval.wd25$kurt[i]=kurtosis(log(tempdat))
  ct=chisq.test(tempdat, p=modDat$meanHarvR[modDat$treat=="wd25"], rescale.p = T)
  pval.wd25$x2[i]=ct$p.value
  f.test=var.test(log(tempdat),log(modDat$meanHarvR[modDat$treat=="wd25"]))
  pval.wd25$f[i]=f.test$statistic
  cis=quantile(log(tempdat), probs=c(0.025, 0.975))
  pval.wd25$lcl[i]=cis[1]
  pval.wd25$ucl[i]=cis[2]
}

# now calculate the number of times the test stat exceeds that of the real data

pval.wd25$sdExceed=0
pval.wd25$medExceed=0
pval.wd25$kurtExceed=0
pval.wd25$x2SigDiff=0
pval.wd25$fStatExceed=0
pval.wd25$coverageExceed=0

wd25.sd=sd(log(modDat$meanHarvR[modDat$treat=="wd25"]))
wd25.med=median(log(modDat$meanHarvR[modDat$treat=="wd25"]))
wd25.kurt=kurtosis(log(modDat$meanHarvR[modDat$treat=="wd25"]))
wd25.x2=chisq.test(modDat$meanHarvR[modDat$treat=="wd25"],p=modDat$meanHarvR[modDat$treat=="wd25"],rescale.p = T)
wd25.f=var.test(log(modDat$meanHarvR[modDat$treat=="wd25"]),log(modDat$meanHarvR[modDat$treat=="wd25"]))

pval.wd25$sdExceed[pval.wd25$sd>wd25.sd]=1
pval.wd25$medExceed[pval.wd25$med>wd25.med]=1
pval.wd25$kurtExceed[pval.wd25$kurt>wd25.kurt]=1 # kurt value great that observed data means there are more outliers, or thicker tails than the observed data
pval.wd25$x2SigDiff[pval.wd25$x2<0.05]=1 # looking at how often we would say the simmed data did not come from the probability distribution of the wd25 data
pval.wd25$fStatExceed[pval.wd25$f>wd25.f$statistic]=1 # how often the f statistic is bigger than the wd25
pval.wd25$coverageExceed[pval.wd25$lcl<=mean(log(modDat$meanHarvR[modDat$treat=="wd25"])) & pval.wd25$ucl>=mean(log(modDat$meanHarvR[modDat$treat=="wd25"]))]=1 # how often the cis from the temp data contain the mean of the wd25 data

#### Pb WD50 ####
pval.wd50=data.frame(alpha=rep(NA,nrow(pars.wd50)),
                     beta=NA,
                     sd=NA,
                     med=NA,
                     kurt=NA,
                     x2=NA,
                     f=NA,
                     lcl=NA,
                     ucl=NA)
set.seed(10)
for(i in 1:nrow(pars.wd50)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="wd50"]),
                 meanlog = pars.wd50[i,1],
                 sdlog = pars.wd50[i,2])
  pval.wd50$alpha[i]=pars.wd50[i,1]
  pval.wd50$beta[i]=pars.wd50[i,2]
  pval.wd50$sd[i]=sd(log(tempdat))
  pval.wd50$med[i]=median(log(tempdat))
  pval.wd50$kurt[i]=kurtosis(log(tempdat))
  ct=chisq.test(tempdat, p=modDat$meanHarvR[modDat$treat=="wd50"], rescale.p = T)
  pval.wd50$x2[i]=ct$p.value
  f.test=var.test(log(tempdat),log(modDat$meanHarvR[modDat$treat=="wd50"]))
  pval.wd50$f[i]=f.test$statistic
  cis=quantile(log(tempdat), probs=c(0.025, 0.975))
  pval.wd50$lcl[i]=cis[1]
  pval.wd50$ucl[i]=cis[2]
}

# now calculate the number of times the test stat exceeds that of the real data

pval.wd50$sdExceed=0
pval.wd50$medExceed=0
pval.wd50$kurtExceed=0
pval.wd50$x2SigDiff=0
pval.wd50$fStatExceed=0
pval.wd50$coverageExceed=0

wd50.sd=sd(log(modDat$meanHarvR[modDat$treat=="wd50"]))
wd50.med=median(log(modDat$meanHarvR[modDat$treat=="wd50"]))
wd50.kurt=kurtosis(log(modDat$meanHarvR[modDat$treat=="wd50"]))
wd50.x2=chisq.test(modDat$meanHarvR[modDat$treat=="wd50"],p=modDat$meanHarvR[modDat$treat=="wd50"],rescale.p = T)
wd50.f=var.test(log(modDat$meanHarvR[modDat$treat=="wd50"]),log(modDat$meanHarvR[modDat$treat=="wd50"]))

pval.wd50$sdExceed[pval.wd50$sd>wd50.sd]=1
pval.wd50$medExceed[pval.wd50$med>wd50.med]=1
pval.wd50$kurtExceed[pval.wd50$kurt>wd50.kurt]=1 # kurt value great that observed data means there are more outliers, or thicker tails than the observed data
pval.wd50$x2SigDiff[pval.wd50$x2<0.05]=1 # looking at how often we would say the simmed data did not come from the probability distribution of the wd50 data
pval.wd50$fStatExceed[pval.wd50$f>wd50.f$statistic]=1 # how often the f statistic is bigger than the wd50
pval.wd50$coverageExceed[pval.wd50$lcl<=mean(log(modDat$meanHarvR[modDat$treat=="wd50"])) & pval.wd50$ucl>=mean(log(modDat$meanHarvR[modDat$treat=="wd50"]))]=1 # how often the cis from the temp data contain the mean of the wd50 data

#### Pb WE50 ####
pval.we50=data.frame(alpha=rep(NA,nrow(pars.we50)),
                     beta=NA,
                     sd=NA,
                     med=NA,
                     kurt=NA,
                     x2=NA,
                     f=NA,
                     lcl=NA,
                     ucl=NA)
set.seed(10)
for(i in 1:nrow(pars.we50)){
  tempdat=rlnorm(n=length(modDat$meanHarvR[modDat$treat=="we50"]),
                 meanlog = pars.we50[i,1],
                 sdlog = pars.we50[i,2])
  pval.we50$alpha[i]=pars.we50[i,1]
  pval.we50$beta[i]=pars.we50[i,2]
  pval.we50$sd[i]=sd(log(tempdat))
  pval.we50$med[i]=median(log(tempdat))
  pval.we50$kurt[i]=kurtosis(log(tempdat))
  ct=chisq.test(tempdat, p=modDat$meanHarvR[modDat$treat=="we50"], rescale.p = T)
  pval.we50$x2[i]=ct$p.value
  f.test=var.test(log(tempdat),log(modDat$meanHarvR[modDat$treat=="we50"]))
  pval.we50$f[i]=f.test$statistic
  cis=quantile(log(tempdat), probs=c(0.025, 0.975))
  pval.we50$lcl[i]=cis[1]
  pval.we50$ucl[i]=cis[2]
}

# now calculate the number of times the test stat exceeds that of the real data

pval.we50$sdExceed=0
pval.we50$medExceed=0
pval.we50$kurtExceed=0
pval.we50$x2SigDiff=0
pval.we50$fStatExceed=0
pval.we50$coverageExceed=0

we50.sd=sd(log(modDat$meanHarvR[modDat$treat=="we50"]))
we50.med=median(log(modDat$meanHarvR[modDat$treat=="we50"]))
we50.kurt=kurtosis(log(modDat$meanHarvR[modDat$treat=="we50"]))
we50.x2=chisq.test(modDat$meanHarvR[modDat$treat=="we50"],p=modDat$meanHarvR[modDat$treat=="we50"],rescale.p = T)
we50.f=var.test(log(modDat$meanHarvR[modDat$treat=="we50"]),log(modDat$meanHarvR[modDat$treat=="we50"]))

pval.we50$sdExceed[pval.we50$sd>we50.sd]=1
pval.we50$medExceed[pval.we50$med>we50.med]=1
pval.we50$kurtExceed[pval.we50$kurt>we50.kurt]=1 # kurt value great that observed data means there are more outliers, or thicker tails than the observed data
pval.we50$x2SigDiff[pval.we50$x2<0.05]=1 # looking at how often we would say the simmed data did not come from the probability distribution of the we50 data
pval.we50$fStatExceed[pval.we50$f>we50.f$statistic]=1 # how often the f statistic is bigger than the we50
pval.we50$coverageExceed[pval.we50$lcl<=mean(log(modDat$meanHarvR[modDat$treat=="we50"])) & pval.we50$ucl>=mean(log(modDat$meanHarvR[modDat$treat=="we50"]))]=1 # how often the cis from the temp data contain the mean of the we50 data


# the following will all be 1 if all parm values in the chain produce data with a CI that covers the mean from the actual creel data.
# (sum(pval.actual$coverageExceed)/nrow(pval.actual))*100
# (sum(pval.noWinter$coverageExceed)/nrow(pval.noWinter))*100
# (sum(pval.mayAug$coverageExceed)/nrow(pval.mayAug))*100
# (sum(pval.wd25$coverageExceed)/nrow(pval.wd25))*100
# (sum(pval.wd50$coverageExceed)/nrow(pval.wd50))*100
# (sum(pval.we50$coverageExceed)/nrow(pval.we50))*100
```

```{r bpValCalcPLOT, fig.cap="Comparison of multiple metrics for calculating Bayesian p-values. Each panel is a different data scenario. The comparisons in each panel are comparing the model simulated data to the observed data for that same sencario, NOT comparing model simulated data from a reduction scenario to the actual data. This comparison to self is to understand whether the model is adequately representing the data and whether the choice of statistic influences the answer to that question. Values close to 0.5 are ideal, values more extreme than 0.1 or 0.9 are considered evidence for a poor model fit (i.e. significant difference betweek model simulated data and the observed data)."}
### summarizing pvalue output
# reminder this is all comparison to self

# making a list object with all the pval objects
self.pvals=list(pval.actual, pval.noWinter, pval.mayAug, pval.wd25, pval.wd50, pval.we50)

all.self.pvals=data.frame(dataSet=c(rep("actual",5),rep("noWinter",5),rep("mayAug",5),rep("wd25",5),rep("wd50",5),rep("we50",5)),
                          measure=rep(c("sd","med","kurt","x2Sig","f"),6),
                          pval=NA)

for(i in 1:length(self.pvals)){
  tp=self.pvals[[i]]
  datSet=unique(all.self.pvals$dataSet)[i]
  
  all.self.pvals$pval[all.self.pvals$dataSet==datSet & all.self.pvals$measure=="cv"]=sum(tp$cvExceed==1)/nrow(tp)
  all.self.pvals$pval[all.self.pvals$dataSet==datSet & all.self.pvals$measure=="sd"]=sum(tp$sdExceed==1)/nrow(tp)
  all.self.pvals$pval[all.self.pvals$dataSet==datSet & all.self.pvals$measure=="med"]=sum(tp$medExceed==1)/nrow(tp)
  all.self.pvals$pval[all.self.pvals$dataSet==datSet & all.self.pvals$measure=="kurt"]=sum(tp$kurtExceed==1)/nrow(tp)
  all.self.pvals$pval[all.self.pvals$dataSet==datSet & all.self.pvals$measure=="x2Sig"]=sum(tp$x2SigDiff==1)/nrow(tp)
  all.self.pvals$pval[all.self.pvals$dataSet==datSet & all.self.pvals$measure=="f"]=sum(tp$fStatExceed==1)/nrow(tp)
}

# not including x2 for this plot since I'm using the calculation differently. Shows kurtosis and median as the only pval metrics that are markedly different from the rest
ggplot(all.self.pvals[all.self.pvals$measure!="x2Sig",])+theme_classic()+
  geom_point(aes(x=measure, y=pval), size=2)+facet_wrap(~dataSet)+
  geom_hline(yintercept = c(0.1,0.9), linetype=2)+
  geom_hline(yintercept = 0.5, linetype=4)+
  labs(y="Bayesian p-value", x="Metric")

```

### 

Figure \@ref(fig:bpValCalcPLOT) shows the p-values resulting from each metric for each data scenario.
Aside from kurtosis and the median, the p-values are quite similar across metrics withing a specific data scenario.

Finally, the choice of 0.1 and 0.9 as the cutoff values for significance here do not seem to influence the outcomes.
A narrower range could be chosen and, unless it were extremely narrow (e.g. 0.4-0.6), the p-values here would all still indicate our models are accurately representing the observed data for all measures except kurtosis and median.
